---
title: "cell size"
author: "DongyueXie"
date: "2020-08-16"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Explore the difference of cell sizes of UMI/non-UMI methods.  

Xin: rpkm and raw counts available

segerstope: rpkm and raw counts available

muraro: Transcript counts were then adjusted to the expected number of molecules based on counts, 256 possible UMI’s and poissonian counting statistics.

baron: raw counts

```{r}
library(MuSiC)
library(xbioc)
library(MASS)
source('code/deconference_main.R')
XinT2Deset <- readRDS("data/MuSiC/XinT2Deset.rds")
xin <- readRDS("data/GSE81608/xin.rds")
segerstolpe <- readRDS("data/MuSiC/segerstolpe.rds")
muraro <- readRDS("data/muraro.rds")
baron <- readRDS("~/deconference/data/baron-human.rds")
GSE50244bulkeset <- readRDS("data/MuSiC/GSE50244bulkeset.rds")
# common genes
common_genes = intersect(intersect(rownames(exprs(XinT2Deset)), rownames(segerstolpe)), intersect(unlist(lapply(strsplit(rownames(muraro),split='__'),function(z){z[1]})), rownames(baron)))

common_genes = (intersect(rownames(GSE50244bulkeset),common_genes))
common_cells = c('alpha','beta','delta','gamma')

length(common_genes)
```

baron data set

```{r}
baron_gene = match(common_genes,rownames(baron))
baron_cell = which(baron$cell_type1%in%common_cells)
out_baron = scRef_multi_proc(counts(baron)[baron_gene,baron_cell],baron$cell_type1[baron_cell],baron$human[baron_cell],estimator='separate',eps=0,
                            est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',meta_mode = "universal",cell_types = common_cells)

out_baron$S_glm
out_baron$S
```

muraro data set

```{r}
muraro_gene =  match(common_genes,unlist(lapply(strsplit(rownames(muraro),split='__'),function(z){z[1]})))
muraro_cell = which(muraro$cell_type1%in%common_cells)
out_muraro = scRef_multi_proc(normcounts(muraro)[muraro_gene,muraro_cell],muraro$cell_type1[muraro_cell],muraro$donor[muraro_cell],estimator='separate',eps=0,
                            est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',meta_mode = "universal",cell_types = common_cells)

out_muraro$S_glm
out_muraro$S
```

Xin data set

Use raw counts:

```{r}
xin_gene =  match(common_genes,rownames(exprs(XinT2Deset)))
xin_cell = which(XinT2Deset$cellType%in%common_cells)
out_xin = scRef_multi_proc(exprs(XinT2Deset)[xin_gene,xin_cell],XinT2Deset$cellType[xin_cell],XinT2Deset$SubjectName[xin_cell],estimator='separate',eps=0,
                            est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',meta_mode = "universal",cell_types = common_cells)

out_xin$S_glm
out_xin$S
```


Use raw counts adjusted by gene length:

```{r}
xin_gene2 = match(common_genes,rownames(normcounts(xin)))
xin_cell2 = which(xin$cell_type1%in%common_cells)
out_xin2 = scRef_multi_proc(t(t(normcounts(xin)[xin_gene,xin_cell]*1e-9)*c(colSums(exprs(XinT2Deset)[xin_gene,xin_cell]))),
                            xin$cell_type1[xin_cell2],xin$donor.id[xin_cell2],estimator='separate',eps=0,
                            est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',meta_mode = "universal",cell_types = common_cells)

out_xin2$S_glm
out_xin2$S
```


segerstolpe data raw counts:

```{r}
segerstolpe_gene =  match(common_genes,unlist(lapply(strsplit(rownames(segerstolpe),split='__'),function(z){z[1]})))
segerstolpe_cell = which(segerstolpe$cell_type1%in%common_cells)
out_segerstolpe = scRef_multi_proc(counts(segerstolpe)[segerstolpe_gene,segerstolpe_cell],
                                   segerstolpe$cell_type1[segerstolpe_cell],segerstolpe$age[segerstolpe_cell],estimator='separate',eps=0,
                                   est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',meta_mode = "universal",cell_types = common_cells)

out_segerstolpe$S_glm
out_segerstolpe$S
```


## combine

combine baron, muraro, xin2

```{r}
S_mat = rbind(out_baron$S_mat,out_muraro$S_mat,out_xin2$S_mat)

S = colMeans(S_mat,na.rm = TRUE)
S = S/S[1]

  # ## method 2: glm
  #
  S_mat_dataframe = data.frame(y = c(S_mat),
                               indi = as.factor(rep(1:nrow(S_mat),ncol(S_mat))),
                               type = as.factor(rep(common_cells,each = nrow(S_mat))))
  fit = glm.nb(y~.,S_mat_dataframe)
  S_glm = exp(fit$coefficients[-c(1:nrow(S_mat))])
  S_glm = c(1,S_glm)
  names(S_glm) = common_cells
  
  S
  
  S_glm
```
