---
title: "cell size"
author: "DongyueXie"
date: "2020-08-16"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Explore the difference of cell sizes of UMI/non-UMI methods.  

Xin: rpkm and raw counts available

segerstope: rpkm and raw counts available

muraro: Transcript counts were then adjusted to the expected number of molecules based on counts, 256 possible UMI’s and poissonian counting statistics.

baron: raw counts

Enge: smart-seq2, raw counts avaiable



```{r}
library(MuSiC)
library(xbioc)
library(MASS)
source('code/deconference_main.R')
xin_raw <- readRDS("data/GSE81608Xin/xin_raw.rds")
xin_rpkm <- readRDS("data/GSE81608Xin/xin_rpkm.rds")

segerstolpe_raw <- readRDS("data/EMTAB5061Segerstolpe/segerstolpe_raw.rds")
segerstolpe_rpkm <- readRDS("data/EMTAB5061Segerstolpe/segerstolpe_rpkm.rds")

enge_raw <- readRDS('data/GSE81547Enge/gse/enge_raw.rds')

muraro <- readRDS("data/GSE85241Muraro/muraro_expected.rds")
baron <- readRDS("data/GSE84133Baron/baron_raw.rds")

GSE50244bulkeset <- readRDS("data/MuSiC/GSE50244bulkeset.rds")

genes <- read.csv("data/GSE81547Enge/hca/genes.csv")
# common genes
common_genes = intersect(intersect(rownames(exprs(xin_raw)), rownames(segerstolpe_raw)), intersect(unlist(lapply(strsplit(rownames(muraro),split='__'),function(z){z[1]})), rownames(baron)))

common_genes = intersect(rownames(enge_raw),common_genes)
common_genes = (intersect(rownames(GSE50244bulkeset),common_genes))
common_genes = intersect(common_genes,genes$featurename)
common_cells = c('alpha','acinar','beta','delta','ductal','gamma')

length(common_genes)

gene_length = (genes$featureend-genes$featurestart)[match(common_genes,genes$featurename)]
```

## baron data set

```{r}
baron_gene = match(common_genes,rownames(baron))
baron_cell = which(baron$cell_type1%in%common_cells)
out_baron = scRef_multi_proc(counts(baron)[baron_gene,baron_cell],baron$cell_type1[baron_cell],baron$human[baron_cell],estimator='separate',eps=0,est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',meta_mode = "universal",cell_types = common_cells)

round(out_baron$S_glm,2)
round(out_baron$S,2)
```

## muraro data set

```{r}
muraro_gene =  match(common_genes,unlist(lapply(strsplit(rownames(muraro),split='__'),function(z){z[1]})))
muraro_cell = which(muraro$cell_type1%in%common_cells)
out_muraro = scRef_multi_proc(normcounts(muraro)[muraro_gene,muraro_cell],muraro$cell_type1[muraro_cell],muraro$donor[muraro_cell],estimator='separate',eps=0,est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',meta_mode = "universal",cell_types = common_cells)

round(out_muraro$S_glm,2)
round(out_muraro$S,2)
```

## Xin data set

### Use raw counts:

```{r}
xin_gene =  match(common_genes,rownames(exprs(xin_raw)))
xin_cell = which(xin_raw$cellType%in%common_cells)
out_xin = scRef_multi_proc(exprs(xin_raw)[xin_gene,xin_cell],xin_raw$cellType[xin_cell],xin_raw$SubjectName[xin_cell],estimator='separate',eps=0,est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',meta_mode = "universal",cell_types = common_cells)

round(out_xin$S_glm,2)
round(out_xin$S,2)
```


### Use raw counts adjusted by gene length:

```{r}
xin_gene2 = match(common_genes,rownames(normcounts(xin_rpkm)))
xin_cell2 = which(xin_rpkm$cell_type1%in%common_cells)
out_xin2 = scRef_multi_proc(t(t(normcounts(xin_rpkm)[xin_gene2,xin_cell2]*1e-6)*c(colSums(exprs(xin_raw)[xin_gene,xin_cell]))),
                            xin_rpkm$cell_type1[xin_cell2],xin_rpkm$donor.id[xin_cell2],estimator='separate',eps=0,
                            est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',meta_mode = "universal",cell_types = common_cells)

round(out_xin2$S_glm,2)
round(out_xin2$S,2)
```

### Use raw counts adjusted by gene length - external gene length data

```{r}

out_xin3 = scRef_multi_proc(exprs(xin_raw)[xin_gene,xin_cell]/gene_length*1e3,
                            xin_raw$cellType[xin_cell],xin_raw$SubjectName[xin_cell],
                            estimator='separate',eps=0,est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',
                            meta_mode = "universal",cell_types = common_cells)

round(out_xin3$S_glm,2)
round(out_xin3$S,2)
```

## segerstolpe data 

### raw counts:

```{r}
segerstolpe_raw_gene =  match(common_genes,unlist(lapply(strsplit(rownames(segerstolpe_raw),split='__'),function(z){z[1]})))
segerstolpe_raw_cell = which(segerstolpe_raw$cell_type1%in%common_cells)
out_segerstolpe_raw = scRef_multi_proc(counts(segerstolpe_raw)[segerstolpe_raw_gene,segerstolpe_raw_cell],
                            segerstolpe_raw$cell_type1[segerstolpe_raw_cell],segerstolpe_raw$age[segerstolpe_raw_cell],
                                   estimator='separate',eps=0,
                                   est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',meta_mode = "universal",cell_types = common_cells)

round(out_segerstolpe_raw$S_glm,2)
round(out_segerstolpe_raw$S,2)
```

### Use raw counts adjusted by gene length:

```{r}
out_segerstolpe2 = scRef_multi_proc(t(t(normcounts(segerstolpe_rpkm)[segerstolpe_raw_gene,segerstolpe_raw_cell]*1e-6)*c(colSums(counts(segerstolpe_raw)[segerstolpe_raw_gene,segerstolpe_raw_cell]))),
                                       segerstolpe_rpkm$cell_type1[segerstolpe_raw_cell],
                                       segerstolpe_rpkm$age[segerstolpe_raw_cell],estimator='separate',eps=0,
                                    est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',meta_mode = "universal",cell_types = common_cells)

round(out_segerstolpe2$S_glm,2)
round(out_segerstolpe2$S,2)
```

### Use raw counts adjusted by gene length - external gene length data

```{r}
out_segerstolpe3 = scRef_multi_proc(counts(segerstolpe_raw)[segerstolpe_raw_gene,segerstolpe_raw_cell]/gene_length*1e3,
                                   segerstolpe_raw$cell_type1[segerstolpe_raw_cell],segerstolpe_raw$age[segerstolpe_raw_cell],
                                   estimator='separate',eps=0,
                                   est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',meta_mode = "universal",cell_types = common_cells)

round(out_segerstolpe3$S_glm,2)
round(out_segerstolpe3$S,2)
```

## Enge data set

### raw

```{r}
enge_gene = match(common_genes,rownames(enge_raw))
enge_cell = which(enge_raw$inferred_cell_type.ch1%in%common_cells)
out_enge_raw = scRef_multi_proc(counts(enge_raw)[enge_gene,enge_cell],enge_raw$inferred_cell_type.ch1[enge_cell],enge_raw$characteristics_ch1[enge_cell],estimator='separate',eps=0,est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',meta_mode = "universal",cell_types = common_cells)
round(out_enge_raw$S_glm,2)
round(out_enge_raw$S,2)
```

### Use raw counts adjusted by gene length - external gene length data

```{r}
out_enge2 = scRef_multi_proc(counts(enge_raw)[enge_gene,enge_cell]/gene_length*1e3,enge_raw$inferred_cell_type.ch1[enge_cell],enge_raw$characteristics_ch1[enge_cell],estimator='separate',eps=0,est_sigma2=F,sigma2=NULL,tau2=NULL,meta_var='plug_in',meta_mode = "universal",cell_types = common_cells)
round(out_enge2$S_glm,2)
round(out_enge2$S,2)
```

## Compare 

```{r}
S_all = rbind(out_baron$S,out_muraro$S,
              out_xin$S,out_xin2$S,out_xin3$S,
              out_segerstolpe_raw$S,out_segerstolpe2$S,out_segerstolpe3$S,
              out_enge_raw$S,out_enge2$S)
rownames(S_all) = c('baron','muraro',
                    'xin_raw','xin/genelen-rpkm','xin/genelen-ext',
                    'segerstolpe_raw','segerstolpe/genelen-rpkm','segerstolpe/genelen-ext',
                    'enge_raw','enge/genelen-ext')
S_all = round(S_all,2)
knitr::kable(S_all,caption = 'ols estimates')
```


```{r}
S_all = rbind(out_baron$S_glm,out_muraro$S_glm,
              out_xin$S_glm,out_xin2$S_glm,out_xin3$S_glm,
              out_segerstolpe_raw$S_glm,out_segerstolpe2$S_glm,out_segerstolpe3$S_glm,
              out_enge_raw$S_glm,out_enge2$S_glm)
rownames(S_all) = c('baron','muraro',
                    'xin_raw','xin/genelen-rpkm','xin/genelen-ext',
                    'segerstolpe_raw','segerstolpe/genelen-rpkm','segerstolpe/genelen-ext',
                    'enge_raw','enge/genelen-ext')
S_all = round(S_all,2)
knitr::kable(S_all,caption = 'glm estimates')
```


## combine

```{r,eval=FALSE}
S_mat = rbind(out_baron$S_mat,out_muraro$S_mat,out_xin2$S_mat)

S = colMeans(S_mat,na.rm = TRUE)
S = S/S[1]

  # ## method 2: glm
  #
  S_mat_dataframe = data.frame(y = c(S_mat),
                               indi = as.factor(rep(1:nrow(S_mat),ncol(S_mat))),
                               type = as.factor(rep(common_cells,each = nrow(S_mat))))
  fit = glm.nb(y~.,S_mat_dataframe)
  S_glm = exp(fit$coefficients[-c(1:nrow(S_mat))])
  S_glm = c(1,S_glm)
  names(S_glm) = common_cells
  
  S
  
  S_glm
```
