---
title: "cell size, remove top expressed genes 0.99"
author: "DongyueXie"
date: "2020-08-25"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE)
```

## Introduction

Explore the difference of cell sizes of UMI/non-UMI methods.  

Xin: rpkm and raw counts available

segerstope: rpkm and raw counts available

muraro: Transcript counts were then adjusted to the expected number of molecules based on counts, 256 possible UMI’s and poissonian counting statistics.

baron: raw counts

Enge: smart-seq2, raw counts avaiable



```{r}
library(MuSiC)
library(xbioc)
library(MASS)
source('code/deconference_main.R')
xin_raw <- readRDS("data/GSE81608Xin/xin_raw.rds")
xin_rpkm <- readRDS("data/GSE81608Xin/xin_rpkm.rds")

segerstolpe_raw <- readRDS("data/EMTAB5061Segerstolpe/segerstolpe_raw.rds")
segerstolpe_rpkm <- readRDS("data/EMTAB5061Segerstolpe/segerstolpe_rpkm.rds")

enge_raw <- readRDS('data/GSE81547Enge/gse/enge_raw.rds')

muraro <- readRDS("data/GSE85241Muraro/muraro_expected.rds")
baron <- readRDS("data/GSE84133Baron/baron_raw.rds")

GSE50244bulkeset <- readRDS("data/MuSiC/GSE50244bulkeset.rds")

genes <- read.csv("data/GSE81547Enge/hca/genes.csv")
# common genes
common_genes = intersect(intersect(rownames(exprs(xin_raw)), rownames(segerstolpe_raw)), intersect(unlist(lapply(strsplit(rownames(muraro),split='__'),function(z){z[1]})), rownames(baron)))

common_genes = intersect(rownames(enge_raw),common_genes)
common_genes = (intersect(rownames(GSE50244bulkeset),common_genes))
common_genes = intersect(common_genes,genes$featurename)
common_cells = c('alpha','acinar','beta','delta','ductal','gamma')

length(common_genes)

gene_length = (genes$featureend-genes$featurestart)[match(common_genes,genes$featurename)]

plott = function(out){
  par(mfrow=c(2,3))
  for(k in 1:6){
    if(!is.nan(out$S[k])){
      
      plot(out$X[,k]*1e3,out$Sigma[,k]*1e6,
       pch=20,col='grey60',
       xlab='mean',ylab='var',
       main=paste(common_cells[k],"cell : mean-var plot"))
  loess_fit = loess(y~.,
                    data.frame(x=out$X[,k]*1e3,y=out$Sigma[,k]*1e6))
  x_seq = seq(range(out$X[,k]*1e3)[1],range(out$X[,k]*1e3)[2],length.out = 1e3)
  lines(x_seq,predict(loess_fit,data.frame(x=x_seq)))
  
    }
  
}
}

```

## baron data set

```{r}
baron_gene = match(common_genes,rownames(baron))
baron_data = set_data_decon(rep(1,length(baron_gene)),counts(baron)[baron_gene,],
                            cell_type_idx=baron$cell_type1,
                            indi_idx = baron$human,
                            cell_types=common_cells)


out_baron = scRef_multi_proc(baron_data$Y,baron_data$cell_type_idx,baron_data$indi_idx,
                             est_sigma2=F,cell_types = baron_data$cell_types,diag_cov=TRUE)

round(out_baron$S_glm,2)
round(out_baron$S,2)
plott(out_baron)
 
```

## muraro data set

```{r}
muraro_gene =  match(common_genes,unlist(lapply(strsplit(rownames(muraro),split='__'),function(z){z[1]})))
muraro_data = set_data_decon(rep(1,length(muraro_gene)),normcounts(muraro)[muraro_gene,],
                            cell_type_idx=muraro$cell_type1,
                            indi_idx = muraro$donor,
                            cell_types=common_cells,max_count_quantile = 0.99)


out_muraro = scRef_multi_proc(muraro_data$Y,muraro_data$cell_type_idx,muraro_data$indi_idx,
                             est_sigma2=F,cell_types = muraro_data$cell_types,diag_cov=TRUE)
round(out_muraro$S_glm,2)
round(out_muraro$S,2)
 
```

## Xin data set

### Use raw counts:

```{r}
xin_gene =  match(common_genes,rownames(exprs(xin_raw)))
xin_cell = which(xin_raw$cellType%in%common_cells)
datax = set_data_decon(rep(1,length(xin_gene)),exprs(xin_raw)[xin_gene,],
                          cell_type_idx = xin_raw$cellType,
                          indi_idx = xin_raw$SubjectName,cell_types = common_cells)
out_xin = scRef_multi_proc(datax$Y,datax$cell_type_idx,datax$indi_idx,
                             est_sigma2=F,cell_types = datax$cell_types,diag_cov=TRUE)

round(out_xin$S_glm,2)
round(out_xin$S,2)

```


### Use raw counts adjusted by gene length:

```{r}
xin_gene2 = match(common_genes,rownames(normcounts(xin_rpkm)))
xin_cell2 = which(xin_rpkm$cell_type1%in%common_cells)
datax = set_data_decon(rep(1,length(xin_gene)),
                       t(t(normcounts(xin_rpkm)[xin_gene2,xin_cell2]*1e-6)*c(colSums(exprs(xin_raw)[xin_gene,xin_cell]))),
                          cell_type_idx = xin_rpkm$cell_type1[xin_cell2],
                          indi_idx = xin_rpkm$donor.id[xin_cell2],cell_types = common_cells)
out_xin2 = scRef_multi_proc(datax$Y,datax$cell_type_idx,datax$indi_idx,
                             est_sigma2=F,cell_types = datax$cell_types,diag_cov=TRUE)

round(out_xin2$S_glm,2)
round(out_xin2$S,2)
plott(out_xin2)
 
```

### Use raw counts adjusted by gene length - external gene length data

```{r}
datax = set_data_decon(rep(1,length(xin_gene)),exprs(xin_raw)[xin_gene,]/gene_length*1e3,
                          cell_type_idx = xin_raw$cellType,
                          indi_idx = xin_raw$SubjectName,cell_types = common_cells)
out_xin3 = scRef_multi_proc(datax$Y,datax$cell_type_idx,datax$indi_idx,
                             est_sigma2=F,cell_types = datax$cell_types,diag_cov=TRUE)

round(out_xin3$S_glm,2)
round(out_xin3$S,2)
plott(out_xin3)
 
```

## segerstolpe data 

### raw counts:

```{r}
segerstolpe_raw_gene =  match(common_genes,unlist(lapply(strsplit(rownames(segerstolpe_raw),split='__'),function(z){z[1]})))
# segerstolpe_raw_cell = which(segerstolpe_raw$cell_type1%in%common_cells)
datax = set_data_decon(rep(1,length(segerstolpe_raw_gene)),
                       counts(segerstolpe_raw)[segerstolpe_raw_gene,],
                            cell_type_idx=segerstolpe_raw$cell_type1,
                            indi_idx = segerstolpe_raw$age,
                            cell_types=common_cells,max_count_quantile = 0.99)

out_segerstolpe_raw = scRef_multi_proc(datax$Y,datax$cell_type_idx,datax$indi_idx,
                             est_sigma2=F,cell_types = datax$cell_types,diag_cov=TRUE)
round(out_segerstolpe_raw$S_glm,2)
round(out_segerstolpe_raw$S,2)
```

### Use raw counts adjusted by gene length:

```{r}
datax = set_data_decon(rep(1,length(segerstolpe_raw_gene)),
                       t(t(normcounts(segerstolpe_rpkm)[segerstolpe_raw_gene,]*1e-6)*c(colSums(counts(segerstolpe_raw)[segerstolpe_raw_gene,]))),
                            cell_type_idx=segerstolpe_raw$cell_type1,
                            indi_idx = segerstolpe_raw$age,
                            cell_types=common_cells,max_count_quantile = 0.99)


out_segerstolpe2 = scRef_multi_proc(datax$Y,datax$cell_type_idx,datax$indi_idx,
                             est_sigma2=F,cell_types = datax$cell_types,diag_cov=TRUE)

round(out_segerstolpe2$S_glm,2)
round(out_segerstolpe2$S,2)
plott(out_segerstolpe2)
 
```

### Use raw counts adjusted by gene length - external gene length data

```{r}
datax = set_data_decon(rep(1,length(segerstolpe_raw_gene)),
                       counts(segerstolpe_raw)[segerstolpe_raw_gene,]/gene_length*1e3,
                            cell_type_idx=segerstolpe_raw$cell_type1,
                            indi_idx = segerstolpe_raw$age,
                            cell_types=common_cells,max_count_quantile = 0.99)

out_segerstolpe3 = scRef_multi_proc(datax$Y,datax$cell_type_idx,datax$indi_idx,
                             est_sigma2=F,cell_types = datax$cell_types,diag_cov=TRUE)

round(out_segerstolpe3$S_glm,2)
round(out_segerstolpe3$S,2)
plott(out_segerstolpe3)
 
```

## Enge data set

### raw

```{r}
par(mfrow=c(2,3))
for(k in 1:6){
  temp_idx = which(enge_raw$inferred_cell_type.ch1==common_cells[k])
  if(length(temp_idx)!=0){
    plot(rowSums(counts(enge_raw)[,temp_idx]),ylab = 'counts',main=common_cells[k])
  }
}
```

```{r}
enge_gene = match(common_genes,rownames(enge_raw))
#enge_cell = which(enge_raw$inferred_cell_type.ch1%in%common_cells)
datax = set_data_decon(rep(1,length(enge_gene)),
                       counts(enge_raw)[enge_gene,],
                            cell_type_idx=enge_raw$inferred_cell_type.ch1,
                            indi_idx = enge_raw$characteristics_ch1,
                            cell_types=common_cells,max_count_quantile = 0.99)
out_enge_raw = scRef_multi_proc(datax$Y,datax$cell_type_idx,datax$indi_idx,
                             est_sigma2=F,cell_types = datax$cell_types,diag_cov=TRUE)
round(out_enge_raw$S_glm,2)
round(out_enge_raw$S,2)
```

### Use raw counts adjusted by gene length - external gene length data

```{r}
datax = set_data_decon(rep(1,length(enge_gene)),
                       counts(enge_raw)[enge_gene,]/gene_length*1e3,
                            cell_type_idx=enge_raw$inferred_cell_type.ch1,
                            indi_idx = enge_raw$characteristics_ch1,
                            cell_types=common_cells,max_count_quantile = 0.99)
out_enge2 = scRef_multi_proc(datax$Y,datax$cell_type_idx,datax$indi_idx,
                             est_sigma2=F,cell_types = datax$cell_types,diag_cov=TRUE)
round(out_enge2$S_glm,2)
round(out_enge2$S,2)
plott(out_enge2)
 
```

## Compare 

```{r}
S_all = rbind(out_baron$S,out_muraro$S,
              out_xin$S,out_xin2$S,out_xin3$S,
              out_segerstolpe_raw$S,out_segerstolpe2$S,out_segerstolpe3$S,
              out_enge_raw$S,out_enge2$S)
rownames(S_all) = c('baron','muraro',
                    'xin_raw','xin/genelen-rpkm','xin/genelen-ext',
                    'segerstolpe_raw','segerstolpe/genelen-rpkm','segerstolpe/genelen-ext',
                    'enge_raw','enge/genelen-ext')
S_all = round(S_all,2)
knitr::kable(S_all,caption = 'ols estimates')
```


```{r}
S_all = rbind(out_baron$S_glm,out_muraro$S_glm,
              out_xin$S_glm,out_xin2$S_glm,out_xin3$S_glm,
              out_segerstolpe_raw$S_glm,out_segerstolpe2$S_glm,out_segerstolpe3$S_glm,
              out_enge_raw$S_glm,out_enge2$S_glm)
rownames(S_all) = c('baron','muraro',
                    'xin_raw','xin/genelen-rpkm','xin/genelen-ext',
                    'segerstolpe_raw','segerstolpe/genelen-rpkm','segerstolpe/genelen-ext',
                    'enge_raw','enge/genelen-ext')
S_all = round(S_all,2)
knitr::kable(S_all,caption = 'glm estimates')
```


## combine cell size

combine baron, muraro, (xin, seger, enge) - ext

```{r}
S_mat = rbind(out_baron$S_mat,out_muraro$S_mat,out_xin3$S_mat,out_segerstolpe3$S_mat,out_enge2$S_mat)

S = colMeans(S_mat,na.rm = TRUE)
S = S/S[1]


S_mat_2 = rbind(out_baron$S,out_muraro$S,
                out_xin3$S,out_segerstolpe3$S,out_enge2$S)
S_mat_2[is.nan(S_mat_2)] = NA
rownames(S_mat_2) = c('baron','muraro','xin','seger','enge')

  # ## method 2: glm
  #

S_mat_dataframe = data.frame(y = c(S_mat),
                               indi = factor(rep(1:nrow(S_mat),ncol(S_mat))),
                               type = factor(rep(common_cells,each = nrow(S_mat)),levels = common_cells))
  fit = MASS::glm.nb(y~.,S_mat_dataframe)

  S_glm = S
  S_glm[which(!is.nan(S))] = c(1,exp(fit$coefficients[-c(1:nrow(S_mat))]))
  names(S_glm) = common_cells
  
  round(S_mat,2)
  round(S_mat_2,2)
  
  round(apply(S_mat_2,2,mean,na.rm=T),2)
  
  round(S,2)
  
  round(S_glm,2)
```
