---
title: "check gene length"
author: "DongyueXie"
date: "2020-09-23"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

I obtained gene length info from 2 resources: [gencode](https://www.gencodegenes.org/), and [HCA](https://data.humancellatlas.org/explore/projects/cddab57b-6868-4be4-806f-395ed9dd635a/expression-matrices).

Now I want to check which one I should TRUST.

```{r}
library(readr)
gene1 <- read_csv("data/GSE81547Enge/hca/genes.csv")
gene2 = read.table('data/all_genes_location.txt')

## first remove genes that are duplicated. 
## these genes share the same gene name but different ensembly id and on different chromesomes.
## in single cell data, we ususally only have gene name so to avoid using the wrong gene length, we remove these genes fisrt.

gene1_name = gene1$featurename
length(gene1_name)

gene1_name_dup = unique(gene1_name[duplicated(gene1_name)])
gene1_name = gene1_name[-which(gene1_name%in%gene1_name_dup)]
gene2_name = gene2$V4
length(gene2_name)
gene2_name_dup = unique(gene2_name[duplicated(gene2_name)])
gene2_name = gene2_name[-which(gene2_name%in%gene2_name_dup)]

length(gene1_name)
length(gene2_name)
cm_gene = intersect(gene1_name,gene2_name)
length(cm_gene)
gene1_idx = match(cm_gene,gene1$featurename)
gene2_idx = match(cm_gene,gene2$V4)

gene1_length = gene1$featureend[gene1_idx] - gene1$featurestart[gene1_idx]
gene2_length = gene2$V3[gene2_idx] - gene2$V2[gene2_idx]
cor(gene1_length,gene2_length)
gene_diff = abs(gene1_length-gene2_length)
summary(gene_diff)
```


Let's look at some genes that have large difference in gene length.

```{r}
cm_gene[which.max(gene_diff)]
gene1[which(gene1$featurename=="AC034268.2"),]
gene2[which(gene2$V4=="AC034268.2"),]
```

Let's check this gene on [ensembl](https://uswest.ensembl.org/Homo_sapiens/Gene/Summary?db=core;g=ENSG00000285939;r=17:53276760-54365634), which shows that it's region is 53,276,760-54,365,634 on chr17.


```{r}
cm_gene[which.max(gene_diff[-which.max(gene_diff)])]
gene1[which(gene1$featurename=="PTCHD1-AS"),]
gene2[which(gene2$V4=="PTCHD1-AS"),]
```

Let's check this gene on [ensembl](https://uswest.ensembl.org/Homo_sapiens/Gene/Summary?db=core;g=ENSG00000233067;r=X:22191895-22235358;t=ENST00000669979), which shows that it's region is 22,191,895-22,235,358 on chrX.

So it seems that the one from gencode is more reliable. 

now generate gene length file 

```{r,eval=FALSE}
gene_idx = match(gene2_name,gene2$V4)
gene_length = gene2$V3[gene_idx] - gene2$V2[gene_idx]
names(gene_length) = gene2$V4[gene_idx]
saveRDS(gene_length,file='data/gene_length.rds')
```
