---
title: "check if V is diagonal"
author: "DongyueXie"
date: "2020-10-05"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Let $\boldsymbol x_g\in R^K$ denote the population level gene expression, and $\boldsymbol x_{ig}\in R^K$ be the individual level one, with $E(\boldsymbol x_{ig}|\boldsymbol x_{g}) = \boldsymbol x_g$ and $Var(\boldsymbol x_{ig}|\boldsymbol x_{g}) = \Sigma_g$.

Previously, we have treated $\Sigma_g$ as a diagonal matrix. 

Each $\boldsymbol x_{ig}$ is not observable and we have to estimate it. For each individual, we have unbiased estimator $\hat{\boldsymbol x}_{ig}$ available and its variance is $D_{ig}$, a diagonal matrix. It's diagonal because we estimate $\boldsymbol x_{ig}$ one cell type by one cell type. 

Let $\hat{\boldsymbol x}_{g} = \sum_i \frac{1}{N_{indi}} \hat{\boldsymbol x}_{ig}$ then $Var(\hat{\boldsymbol x}_{g}) = \sum_i (\frac{1}{N_{indi}})^2(\Sigma_g+D_{ig})$. To check if $\Sigma_g$ is diagonal, we need to estimate $Var(\hat{\boldsymbol x}_{g})$ and see if it's close to a diagonal matrix. An estimate of $Var(\hat{\boldsymbol x}_{g})$ is $\frac{1}{N_{indi}}\frac{1}{N_{indi}-1}\sum_i(\hat{\boldsymbol x}_{ig} - \hat{\boldsymbol x}_{g})^2$.

[If $\hat{\boldsymbol x}_{i}\sim [{\boldsymbol \mu},V_i]$, one estimator of $\boldsymbol \mu$ is $(\sum_iV_i^{-1/2})^{-1}(\sum_iV_i^{-1/2}\boldsymbol x_i)$. This is derived from scaling each $\boldsymbol x_i$ to have diagonal covariance matrix.]

```{r}
source('code/deconference_main.R')
ref.obj = readRDS('data/pancreas/refs.rds')
cell_types = c('acinar','alpha','beta','delta','ductal','gamma')
temp = getXV_array_all(ref.obj=ref.obj,tau2 = NULL,cell_types=cell_types,
                         indis = NULL,gene_length_adjust=F,gene_length = NULL,
                         protocol = c('UMI','UMI','nonUMI','nonUMI','nonUMI'))

design.mat = getXV(temp$all_X_array,temp$all_Vg_array,
                     S_olss=temp$S_olss,S_glms=temp$S_glms,sigma2=NULL,
                   est_sigma2=F,meta_var=meta_var,meta_mode=meta_mode,
                   cell_types=cell_types,cellsize_est='glm')
```

Let's take a look at the covariance matrix

```{r}
round(cov2cor(matrix(colSums(design.mat$Vg),ncol=6)),3)
```


