---
title: "check coverage after adding weights, Xin data"
author: "DongyueXie"
date: "2021-07-23"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

Possible reasons: 

1. weights depend on data

2. CLT fails: estimates are not normally distributed, variances are under-estimated

3. correlations are not well adjusted.

```{r,warning=F,message=F}
source('code/deconference_main.R')
source('code/simulation/simu_correlation_ult.R')

xin_raw <- readRDS("data/pancreas/xin_raw.rds")
cell_types = c('alpha', 'beta', 'delta', 'gamma')
K = length(cell_types)
rm.indi = c("Non T2D 4","Non T2D 7","Non T2D 10","Non T2D 12")
rm.indi.idx = which(xin_raw$individual%in%rm.indi)

datax.xin = set_data_decon(Y = xin_raw[,-rm.indi.idx],cell_types = cell_types, 
                           gene_thresh = 0.05,max_count_quantile_celltype = 0.95,
                           max_count_quantile_indi = 0.95,
                           w=1)
design.mat.xin = scRef_multi_proc(datax.xin$Y,datax.xin$cell_type_idx,
                                  datax.xin$indi_idx,estimator="separate",
                                  est_sigma2 = TRUE,meta_mode = 'local',smooth.sigma = F)

ref = design.mat.xin$X
sigma2 = design.mat.xin$Sigma



# set.seed(12345)
# ref = matrix(rnorm(G*K),nrow=G)
# ref = abs(ref)
# ref = apply(ref, 2, function(z){z/sum(z)})*G
# sigma2 = ref/2
# rownames(ref) = 1:G



```

```{r}
ref = ref+1/nrow(ref)
sigma2 = sigma2 + 1/nrow(ref)

b1 = c(0.1,0.1,0.3,0.5)
b.m = t(b1%*%t(rep(1,100)))

#'@param re_sample_ref whether re-sample reference at each simulation

#'@param use_all_pair_for_cov  if all_pair_calc_cor, use all pairs for calculating variance
simu_study = function(ref,
                            b,
                            R=NULL,
                            sigma2,
                         
                            nreps = 100,
                            bulk_lib_size = 500,
                            n_indi = 10,
                            
                            verbose=FALSE,
                            alpha=0.05,
                            alpha.cor = 0.1,
                            
                            centeringXY = FALSE,
                            true.beta.for.Sigma=FALSE,
                            calc_cov = F,
                            est_cor = F,
                            only.scale.pos.res = FALSE,
                            only.add.pos.res = FALSE,
                            n_bulk_for_cor = 100,
                            cor_method = 'testing',
                           
                            weighted = FALSE,
                            use.true.sigma.for.weight = FALSE,
                            use.random.weight = FALSE,
                            re_sample_ref = FALSE,
                      use_all_pair_for_cov=FALSE
                            ){

  is.identity = function(X){
    if(!is.null(X)){
      (sum(X)==nrow(X))
    }else{
      FALSE
    }
    
  }

  genp = function(K){
    p = runif(K)
    p/sum(p)
  }

  G = nrow(ref)
  K = ncol(ref)

  
  is.indep = (is.identity(R))|(is.null(R))

 
  

  ## 2. adjust for variance of U_hat, do not adjust for correlation, then hc0, hc2, hc3.
  ## 3. adjust for varaince of U_hat, adjust for correlation, then hc0, hc2, hc3.

  
  gene_names = rownames(ref)
  


  ## pre calculate MLN and generate independent normal

  n.ref = matrix(nrow=G,ncol=K)
  n.Sigma.chol = list()
  n.Sigma = matrix(nrow=G,ncol=K)
  if(!is.indep){
    chol.R = chol(R)
  }
  
  for(k in 1:K){
    n.ref[,k] = log(ref[,k]^2/sqrt(ref[,k]^2+sigma2[,k]))
    n.s = sqrt(log(1+sigma2[,k]/ref[,k]^2))
    if(!is.indep){
      n.Sigma.chol[[k]] = t(n.s*t(chol.R))
    }else{
      n.Sigma[,k] = n.s^2
    }
  }


  if(!est_cor){
    if(!is.indep){
      cor.idx = which(R!=0,arr.ind = T)
      cor.idx = cor.idx[(cor.idx[,1]!=cor.idx[,2]),]
    }else{
      cor.idx = NULL
    }
    
  }else{
  
      X_array = array(dim=c(G,K,n_bulk_for_cor))
      for(k in 1:K){
      if(is.indep){
        X_array[,k,] = exp(matrix(rnorm(G*n_bulk_for_cor,n.ref[,k],sqrt(n.Sigma[,k])),ncol=n_bulk_for_cor))
      }else{
        X_array[,k,] = t(exp(mvnfast::rmvn(n_bulk_for_cor,mu = n.ref[,k],sigma = n.Sigma.chol[[k]],isChol = TRUE)))
      }
      }
      mb = apply(X_array,3,function(z){z%*%genp(K)})
      thetab = apply(mb,2,function(z){z/sum(z)})
      bulk_for_cor = matrix(rpois(G*n_bulk_for_cor,bulk_lib_size*G*thetab),nrow=G)
      rownames(bulk_for_cor) = gene_names
      cor.idx = get_cor_pairs2(bulk_for_cor,alpha=alpha.cor,method=cor_method)
      
      
  }

  if(!re_sample_ref){
    n.temp = n_indi+nreps
    X_array = array(dim=c(G,K,n.temp))

    for(k in 1:K){
      if(is.indep){
        X_array[,k,] = exp(matrix(rnorm(G*n.temp,n.ref[,k],sqrt(n.Sigma[,k])),ncol=n.temp))
      }else{
        X_array[,k,] = t(exp(mvnfast::rmvn(n.temp,mu = n.ref[,k],sigma = n.Sigma.chol[[k]],isChol = TRUE)))
      }
    }

    #browser()

    X_array_bulk = X_array[,,1:nreps]
    b = b%*%t(rep(1,nreps))
    mb = lapply(1:nreps,function(i){X_array_bulk[,,i]%*%b[,i]})
    mb = do.call(cbind,mb)
    true_betas = t(t(b)*c(apply(mb,2,function(z){bulk_lib_size*G/sum(z)})))
    thetab = apply(mb,2,function(z){z/sum(z)})


    #browser()
    y = matrix(rpois(G*nreps,bulk_lib_size*G*thetab),nrow=G)
    rownames(y) = gene_names

    X_array_ref = X_array[,,(nreps+1):(nreps+n_indi)]

    X = apply(X_array_ref,c(1,2),mean,na.rm=TRUE)
    V = t(apply(X_array_ref,c(1),function(z){(cov(t(z),use = 'complete.obs'))}))/n_indi
    
    if(weighted){
      if(use.true.sigma.for.weight){
        w = 1/rowSums(sigma2)
      }else if(use.random.weight){
        w = runif(G)
      }else{
          fit.vash = vashr::vash(sqrt(rowSums(V)),df=n_indi-1)
          w = 1/(fit.vash$sd.post)^2
      }
       # calc weights for each cell type, then average,

    }else{
      w = 1
    }
    
    fit.hc3 = estimation_func2(y=y,X=X,Vg=V,
                                   w=w,hc.type='hc3',correction=FALSE,
                                   calc_cov=calc_cov,verbose=verbose,
                                   cor.idx=cor.idx,
                                   centeringXY=centeringXY,
                                   true.beta = if(true.beta.for.Sigma){true_betas}else{NULL},
                                   only.scale.pos.res=only.scale.pos.res,
                                  only.add.pos.res = only.add.pos.res,
                               use_all_pair_for_cov=use_all_pair_for_cov)
    
    p_hat = t(fit.hc3$beta_hat)
    p_hat_se = t(fit.hc3$beta_se)
    true_betas = t(true_betas)
    
  }else{
    
    n_bulk = 1
  p_hat = matrix(nrow=nreps,ncol=K)
   p_hat_se= matrix(nrow=nreps,ncol=K)
   true_betas = matrix(nrow=nreps,ncol=K)


  for(reps in 1:nreps){
    
    #if(reps%%10==0){print(sprintf("running %d (out of %d)",reps,nreps))}
  
    # generate individual reference matrices


    n.temp = n_indi+n_bulk
    X_array = array(dim=c(G,K,n.temp))

    for(k in 1:K){
      if(is.indep){
        X_array[,k,] = exp(matrix(rnorm(G*n.temp,n.ref[,k],sqrt(n.Sigma[,k])),ncol=n.temp))
      }else{
        X_array[,k,] = t(exp(mvnfast::rmvn(n.temp,mu = n.ref[,k],sigma = n.Sigma.chol[[k]],isChol = TRUE)))
      }
    }

    #browser()

    X_array_bulk = X_array[,,1:n_bulk]

    X_array_ref = X_array[,,(n_bulk+1):(n_bulk+n_indi)]

    



    mb = cbind(X_array_bulk%*%b)
    
    true.beta = t(t(b)*c(apply(mb,2,function(z){bulk_lib_size*G/sum(z)})))
    true_betas[reps,] = c(true.beta)
    thetab = apply(mb,2,function(z){z/sum(z)})


    #browser()
    y = matrix(rpois(G*n_bulk,bulk_lib_size*G*thetab),nrow=G)
    rownames(y) = gene_names

    #bulks = SingleCellExperiment(assays = list(counts = y),
    #                             colData = DataFrame(individual = 1:n_bulk))



    ## generate reference matrices X

    # X_array = array(dim=c(G,K,n_indi))
    #
    # for(k in 1:K){
    #   x_k = sim_MLN(n_indi,ref[,k],sqrt(sigma2[,k]),R)
    #   X_array[,k,] = t(x_k)
    # }


    ## get correlated pairs

    # if(est_cor){
    # 
    #   # generate bulk data for estimating correlation
    #   if(re_sample_for_cor){
    #     
    #     mb = apply(X_array_bulk_for_cor,3,function(z){z%*%genp(K)})
    #     thetab = apply(mb,2,function(z){z/sum(z)})
    #     bulk_for_cor = matrix(rpois(G*n_bulk_for_cor,bulk_lib_size*G*thetab),nrow=G)
    #     rownames(bulk_for_cor) = gene_names
    # 
    #     cor.idx = get_cor_pairs2(bulk_for_cor,alpha=alpha.cor,method=cor_method)
    #   
    #   }
    # 
    # 
    # }

    # fit model

    #browser()

    X = apply(X_array_ref,c(1,2),mean,na.rm=TRUE)
    V = t(apply(X_array_ref,c(1),function(z){(cov(t(z),use = 'complete.obs'))}))/n_indi


    if(weighted){
      if(use.true.sigma.for.weight){
        w = 1/rowSums(sigma2)
      }else if(use.random.weight){
        w = runif(G)
      }else{
          fit.vash = vashr::vash(sqrt(rowSums(V)),df=n_indi-1)
          w = 1/(fit.vash$sd.post)^2
      }
       # calc weights for each cell type, then average,

    }else{
      w = 1
    }

    # fit.hc0 = estimation_func2(y=y,X=X,Vg=V,
    #                                w=w,hc.type='hc0',correction=FALSE,
    #                                calc_cov=calc_cov,verbose=verbose,
    #                                cor.idx=cor.idx,
    #                                centeringXY=centeringXY,
    #                                true.beta = if(true.beta.for.Sigma){true.beta}else{NULL},
    #                                only.scale.pos.res=only.scale.pos.res)

    # fit.adj.hc2 = estimation_func2(y=y,X=X,Vg=V,
    #                                w=w,hc.type='hc2',correction=FALSE,
    #                                calc_cov=calc_cov,verbose=verbose,
    #                                cor.idx=cor.idx,
    #                                centeringXY=centeringXY,
    #                                true.beta = if(true.beta.for.Sigma){true.beta}else{NULL},
    #                                only.scale.pos.res=only.scale.pos.res)

    fit.hc3 = estimation_func2(y=y,X=X,Vg=V,
                                   w=w,hc.type='hc3',correction=FALSE,
                                   calc_cov=calc_cov,verbose=verbose,
                                   cor.idx=cor.idx,
                                   centeringXY=centeringXY,
                                   true.beta = if(true.beta.for.Sigma){true.beta}else{NULL},
                                   only.scale.pos.res=only.scale.pos.res,
                               only.add.pos.res = only.add.pos.res,
                               use_all_pair_for_cov=use_all_pair_for_cov)
    
    p_hat[reps,] = c(fit.hc3$beta_hat)
    p_hat_se[reps,] = c(fit.hc3$beta_se)
    
    #fit.lm = unadjusted_lm(y,X,w)

    # fit.adj.jack = estimation_func2(y=y,X=X,Vg=V,
    #                                w=w,hc.type='jackknife',correction=FALSE,
    #                                calc_cov=calc_cov,verbose=verbose,
    #                                cor.idx=cor.idx,
    #                                centeringXY=centeringXY,
    #                                true.beta = if(true.beta.for.Sigma){true.beta}else{NULL},
    #                                only.scale.pos.res=only.scale.pos.res,
    #                                nfold = nfold)

    # fit.unadj.hc0 = estimation_func2(y=y,X=X,Vg=V,
    #                                  w=w,hc.type='hc0',correction=FALSE,
    #                                  calc_cov=calc_cov,verbose=verbose,
    #                                  cor.idx=NULL,
    #                                  centeringXY=centeringXY,
    #                                  true.beta = if(true.beta.for.Sigma){true.beta}else{NULL},
    #                                  only.scale.pos.res=only.scale.pos.res)
    # 
    # fit.unadj.hc2 = estimation_func2(y=y,X=X,Vg=V,
    #                                  w=w,hc.type='hc2',correction=FALSE,
    #                                  calc_cov=calc_cov,verbose=verbose,
    #                                  cor.idx=NULL,
    #                                  centeringXY=centeringXY,
    #                                  true.beta = if(true.beta.for.Sigma){true.beta}else{NULL},
    #                                  only.scale.pos.res=only.scale.pos.res)
    # 
    # fit.unadj.hc3 = estimation_func2(y=y,X=X,Vg=V,
    #                                  w=w,hc.type='hc3',correction=FALSE,
    #                                  calc_cov=calc_cov,verbose=verbose,
    #                                  cor.idx=NULL,
    #                                  centeringXY=centeringXY,
    #                                  true.beta = if(true.beta.for.Sigma){true.beta}else{NULL},
    #                                  only.scale.pos.res=only.scale.pos.res)
  


  }
  }

   

  return(list(p_hat=p_hat,p_hat_se = p_hat_se,true_betas=true_betas))



}
rmse = function(x,y){sqrt(mean((x-y)^2))}


summary_simu_study = function(out1,b=c(0.1,0.1,0.3,0.5)){
  b.m = t(b%*%t(rep(1,nrow(out1$p_hat))))
  rmses = c()
  for(k in 1:K){
    rmses = c(rmses,rmse(out1$p_hat[,k],b.m[,k]))
  }

  print('rmse')
  print(round(rmses,3))

  print(round(rmse(out1$p_hat,b.m),3))
  
  z.err.hc3 = (out1$p_hat-b.m)/out1$p_hat_se
  print('coverage')
  print(round(colMeans(abs(z.err.hc3)<1.96,na.rm=T),3))
  
  par(mfrow=c(2,2))
  for(k in 1:K){
    qqnorm(out1$p_hat[,k],main = 'qqplot - p_hat')
    qqline(out1$p_hat[,k])
  }
  
  
  for(k in 1:K){
    qqnorm(z.err.hc3[,k],main = 'qqplot - z score')
    qqline(z.err.hc3[,k])
  }
  
  for(k in 1:K){
    h = hist(z.err.hc3[,k],breaks = 20,main='histogram of z score',xlab=paste('p_1',k))
    xfit<-seq(-3,3,length=40)
    yfit<-dnorm(xfit,mean=0,sd=1)
    yfit <- yfit*diff(h$mids[1:2])*length(z.err.hc3[,k])
    lines(xfit, yfit, col="blue", lwd=2)
  }
  par(mfrow=c(1,1))
  
  boxplot(out1$p_hat_se,col='grey80',ylab='se',ylim = range(apply(out1$p_hat_se,2,quantile,c(0,0.95),na.rm=TRUE)))
  lines(apply(out1$p_hat, 2, sd),type='p',col='yellow',pch=17)
  legend('bottomright',c('true sd'),pch=17,col='yellow')
}

```




## No correlation, No weights

### p_hat for variance estimation

```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = NULL,sigma2=sigma2,n_indi=10,weighted = F,use.random.weight = F)
# saveRDS(out1,file='output/check_coverage_weight_xin/uncor_no_weight.rds')
out1 = readRDS('output/check_coverage_weight_xin/uncor_no_weight.rds')
summary_simu_study(out1)
```

### true p for variance estimation

```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = NULL,sigma2=sigma2,n_indi=10,weighted = F,use.random.weight = F,true.beta.for.Sigma = T)
# saveRDS(out1,file='output/check_coverage_weight_xin/uncor_no_weight_truep.rds')
out1 = readRDS('output/check_coverage_weight_xin/uncor_no_weight_truep.rds')
summary_simu_study(out1)
```

### p_hat for variance estimation, use all pairs

```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = NULL,sigma2=sigma2,n_indi=10,weighted = F,use.random.weight = F,use_all_pair_for_cov = T)
# saveRDS(out1,file='output/check_coverage_weight_xin/uncor_no_weight_all_pair.rds')
out1 = readRDS('output/check_coverage_weight_xin/uncor_no_weight_all_pair.rds')
summary_simu_study(out1)
```

### true p for variance estimation

```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = NULL,sigma2=sigma2,n_indi=10,weighted = F,use.random.weight = F,true.beta.for.Sigma = T,use_all_pair_for_cov = T)
# saveRDS(out1,file='output/check_coverage_weight_xin/uncor_no_weight_truep_all_pair.rds')
out1 = readRDS('output/check_coverage_weight_xin/uncor_no_weight_truep_all_pair.rds')
summary_simu_study(out1)
```


## No correlation, random weights


```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = NULL,sigma2=sigma2,n_indi=10,weighted = TRUE,use.random.weight = TRUE)
# saveRDS(out1,file='output/check_coverage_weight_xin/uncor_random_weight.rds')
out1 = readRDS('output/check_coverage_weight_xin/uncor_random_weight.rds')
summary_simu_study(out1)
```

## No correlation, true variance for  weights


```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = NULL,sigma2=sigma2,n_indi=10,weighted = TRUE,use.random.weight = F,use.true.sigma.for.weight = T)
# saveRDS(out1,file='output/check_coverage_weight_xin/uncor_truevar_weight.rds')
out1 = readRDS('output/check_coverage_weight_xin/uncor_truevar_weight.rds')
summary_simu_study(out1)
```

## No correlation, estimated variance as weights

### p hat for variance estimation

```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = NULL,sigma2=sigma2,n_indi=10,weighted = TRUE,
#                   use.random.weight = FALSE,use.true.sigma.for.weight = FALSE)
# saveRDS(out1,file='output/check_coverage_weight_xin/uncor_estvar_weight.rds')
out1 = readRDS('output/check_coverage_weight_xin/uncor_estvar_weight.rds')
summary_simu_study(out1)

```

### true p for variance estimation

```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = NULL,sigma2=sigma2,n_indi=10,weighted = TRUE,
#                   use.random.weight = FALSE,use.true.sigma.for.weight = FALSE,true.beta.for.Sigma = T)
# saveRDS(out1,file='output/check_coverage_weight_xin/uncor_estvar_weight_truep.rds')
out1 = readRDS('output/check_coverage_weight_xin/uncor_estvar_weight_truep.rds')
summary_simu_study(out1)

```

## Correlated, no weights, known correlation



```{r}
G = nrow(ref)
K = 4
d = 500
R = matrix(0,nrow=G,ncol=G)

for(i in 1:G){
  for(j in i:min(i+d,G)){
    R[i,j] = max(1-abs(i-j)/d,0)
  }
}
R = R+t(R) - diag(G)
library(Matrix)
R = Matrix(R,sparse = TRUE)
```

### p_hat for variance

```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = R,sigma2=sigma2,n_indi=10,weighted = F,est_cor = F)
# saveRDS(out1,file='output/check_coverage_weight_xin/cor_no_weight_known_cor.rds')
out1 = readRDS('output/check_coverage_weight_xin/cor_no_weight_known_cor.rds')
summary_simu_study(out1)
```

### true p for variance

```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = R,sigma2=sigma2,n_indi=10,weighted = F,est_cor = F,true.beta.for.Sigma = T)
# saveRDS(out1,file='output/check_coverage_weight_xin/cor_no_weight_known_cor_truep.rds')
out1 = readRDS('output/check_coverage_weight_xin/cor_no_weight_known_cor_truep.rds')
summary_simu_study(out1)
```

## Correlated, random weights, known correlation


```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = R,sigma2=sigma2,n_indi=10,weighted = T,est_cor = F,use.random.weight = TRUE)
# saveRDS(out1,file='output/check_coverage_weight_xin/cor_random_weight_known_cor.rds')
out1 = readRDS('output/check_coverage_weight_xin/cor_random_weight_known_cor.rds')
summary_simu_study(out1)
```


## Correlated, true variance for weights, known correlation

### p_hat for variance

```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = R,sigma2=sigma2,n_indi=10,weighted = T,est_cor = F,use.true.sigma.for.weight = T)
# saveRDS(out1,file='output/check_coverage_weight_xin/cor_truevar_weight_known_cor.rds')
out1 = readRDS('output/check_coverage_weight_xin/cor_truevar_weight_known_cor.rds')
summary_simu_study(out1)
```

### true p for variance

```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = R,sigma2=sigma2,n_indi=10,weighted = T,est_cor = F,use.true.sigma.for.weight = T,true.beta.for.Sigma = T)
# saveRDS(out1,file='output/check_coverage_weight_xin/cor_truevar_weight_known_cor_truep.rds')
out1 = readRDS('output/check_coverage_weight_xin/cor_truevar_weight_known_cor_truep.rds')
summary_simu_study(out1)
```


## Correlated, estimated variance for weights, known correlation

### p_hat for variance

```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = R,sigma2=sigma2,n_indi=10,weighted = T,est_cor = F,
#                   use.random.weight = FALSE,use.true.sigma.for.weight = FALSE)
# saveRDS(out1,file='output/check_coverage_weight_xin/cor_estvar_weight_known_cor.rds')
out1 = readRDS('output/check_coverage_weight_xin/cor_estvar_weight_known_cor.rds')
summary_simu_study(out1)
```

### true p for variance


```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = R,sigma2=sigma2,n_indi=10,weighted = T,est_cor = F,
#                  use.random.weight = FALSE,use.true.sigma.for.weight = FALSE,true.beta.for.Sigma = T)
# saveRDS(out1,file='output/check_coverage_weight_xin/cor_estvar_weight_known_cor_truep.rds')
out1 = readRDS('output/check_coverage_weight_xin/cor_estvar_weight_known_cor_truep.rds')
summary_simu_study(out1)
```

## Correlated, estimated variance for weights, all pairs for variance

### p_hat for variance

```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = R,sigma2=sigma2,n_indi=10,weighted = T,est_cor = F,
#                  use.random.weight = FALSE,use.true.sigma.for.weight = FALSE,use_all_pair_for_cov = TRUE)
# saveRDS(out1,file='output/check_coverage_weight_xin/cor_estvar_weight_all_pair.rds')
out1 = readRDS('output/check_coverage_weight_xin/cor_estvar_weight_all_pair.rds')
summary_simu_study(out1)
```

### true for variance

```{r}
# set.seed(12345)
# out1 = simu_study(ref,b1,R = R,sigma2=sigma2,n_indi=10,weighted = T,est_cor = F,
#                  use.random.weight = FALSE,use.true.sigma.for.weight = FALSE,use_all_pair_for_cov = TRUE,true.beta.for.Sigma = T)
# saveRDS(out1,file='output/check_coverage_weight_xin/cor_estvar_weight_all_pair_truep.rds')
out1 = readRDS('output/check_coverage_weight_xin/cor_estvar_weight_all_pair_truep.rds')
summary_simu_study(out1)
```
