---
title: "Simulation on real data set, with comparisons to MuSiC and CibersortX"
author: "Dongyue Xie"
date: "2021-04-09"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction


See [here](gene_weights.html) for a previous simulation on choosing gene weights.


In MuSiC:

  The default is for bulk data Y: convert y such that it sums to 1 and then multiply y by 100; for reference matrix X: it's the relative expression matrix whose each column sum to 1; it's default mode also multiply X by cell size S, where S is very big if estimated from single cell data.
  
  The above scaling makes the parameter nu in the weight computation important, where $weight = 1/(\nu+r^2+across.indi.var)$. 

```{r}
library(MuSiC)
library(Biobase)
library(xbioc)
source('code/deconference_main.R')
source('code/utils.R')
source('code/wols.R')
seger <- readRDS("data/pancreas/segerstolpe_raw.rds")

#XinT2D.eset <- readRDS("data/MuSiC/XinT2Deset.rds")
#XinT2D.construct.full = bulk_construct(XinT2D.eset, clusters = 'cellType', samples = 'SubjectName')
#XinT2D.construct.full$prop.real = relative.ab(XinT2D.construct.full$num.real, by.col = FALSE)
xin_raw <- readRDS("data/pancreas/xin_raw.rds")
baron = readRDS("data/pancreas/baron.rds")
```


```{r}

bulk_mat_to_es = function(Y){
  ExpressionSet(assayData = Y)
}


#'@param b either a vector or a matrix of dimension K*n_bulk
createBulk_no.tech.noise = function(X_array, b,gene_names){
  if(is.null(dim(b))){
    y = apply(X_array,3,function(z){z%*%b})
  }else{
    G = dim(X_array)[1]
    K = dim(X_array)[2]
    Ni = dim(X_array)[3]
    y = matrix(nrow=G,ncol=Ni)
    for(i in 1:Ni){
      y[,i] = X_array[,,i]%*%b[,i]
    }
  }
  rownames(y) = gene_names
  colnames(y) = 1:ncol(y)
  list(y=y,b=b)
}

createBulk_tech.noise = function(X_array, b, gene_names, ls_per_gene = 500,seed=12345){
  
  G = dim(X_array)[1]
  K = dim(X_array)[2]
  Ni = dim(X_array)[3]
    
  if(is.null(dim(b))){
    y = apply(X_array,3,function(z){z%*%b})
  }else{
    y = matrix(nrow=G,ncol=Ni)
    for(i in 1:Ni){
      y[,i] = X_array[,,i]%*%b[,i]
    }
  }
  y = y/colSums(y)*ls_per_gene*G
  set.seed(seed)
  y = matrix(rpois(prod(dim(y)),y),nrow=nrow(y),ncol=ncol(y))
  rownames(y) = gene_names
  colnames(y) = 1:ncol(y)
  list(y=y,b=b*ls_per_gene)
}


calc_weight = function(){
  ####
}





is.covered = function(b,bhat,bhat.se,alpha=0.05){
  (b<=bhat+qnorm(1-alpha/2)*bhat.se)&(b>=bhat-qnorm(1-alpha/2)*bhat.se)
}

#'@param design.mat output from scRef_multi_proc, in which X is the reference matrix.
#'@param datax output from set_data_decon, in which y is the bulk data matrix.
#'@param b betas, K by n_bulk.


diag.cov = function(X){
  #center X
  X = scale(X,center=TRUE,scale=F)
  colSums(X^2)/(nrow(X)-1)
}


```

## Create bulk data(with technical noise) using Xin dataset.

```{r}
cell_types = c('alpha', 'beta', 'delta', 'gamma')
K = length(cell_types)
rm.indi = c("Non T2D 4","Non T2D 7","Non T2D 10","Non T2D 12")
#rm.indi = levels(XinT2D.eset$SubjectName)[rm.indi]
rm.indi.idx = which(xin_raw$individual%in%rm.indi)
datax.xin = set_data_decon(Y = xin_raw[,-rm.indi.idx],cell_types = cell_types, gene_thresh = 0,max_count_quantile = 1,w=1)
design.mat.xin = scRef_multi_proc(datax.xin$Y,datax.xin$cell_type_idx,datax.xin$indi_idx,estimator="separate",est_sigma2 = FALSE)
# create bulk data
b = c(0.1,0.1,0.3,0.5)
bulk.xin = createBulk_tech.noise(design.mat.xin$X_array,b,gene_names = rownames(xin_raw),seed=12345,ls_per_gene = 500)
```

### Seger as reference

```{r}
# fit music
seger.es = sce_to_es(seger)
cell_size = data.frame(type = c('alpha', 'beta', 'delta', 'gamma'), size=c(1,1,1,1))
fit.music = music_prop(bulk_mat_to_es(bulk.xin$y),seger.es,clusters = 'cell_type', samples = 'individual',
                       select.ct = cell_types,cell_size=cell_size,verbose = FALSE,nu=1e-4,normalize = F)
rmse(fit.music$Est.prop.weighted,(rep(1,14))%*%t(b))
round(fit.music$Est.prop.weighted,2)

fit.music2 = music_prop(bulk_mat_to_es(bulk.xin$y),seger.es,clusters = 'cell_type', samples = 'individual',
                       select.ct = cell_types,cell_size=cell_size,verbose = FALSE,nu=1e-4,normalize = T)
rmse(fit.music2$Est.prop.weighted,(rep(1,14))%*%t(b))
round(fit.music2$Est.prop.weighted,2)

fit.music3 = music_prop(bulk_mat_to_es(bulk.xin$y),seger.es,clusters = 'cell_type', samples = 'individual',
                       select.ct = cell_types,cell_size=cell_size,verbose = FALSE,nu=1e-2,normalize = T)
rmse(fit.music3$Est.prop.weighted,(rep(1,14))%*%t(b))
round(fit.music3$Est.prop.weighted,2)
```


```{r}
datax = set_data_decon(y = bulk.xin$y, Y = seger,
                       cell_types = cell_types, 
                       gene_thresh = 0,max_count_quantile = 1,w=1)

design.mat = scRef_multi_proc(datax$Y,datax$cell_type_idx,datax$indi_idx,
                                    estimator="separate",est_sigma2 = FALSE)

fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc0',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = T)
rmse(fit.my$beta_hat,(rep(1,14))%*%t(b))
round(fit.my$beta_hat,2)
round(fit.my$beta_hat_se,3)
is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))

```


```{r}
diag(cor(fit.my$weight,fit.music$Weight.gene,use = "complete.obs"))
diag(cor(fit.my$X,fit.music$X,use = "pairwise.complete.obs"))
diag(cor(fit.my$Vg[,c(1,6,11,16)],fit.music$Sigma,use = "pairwise.complete.obs"))
```

scale y as y/sum(y)*100

```{r}
fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc0',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = T,use.weight.for.var = T)
rmse(fit.my$beta_hat,(rep(1,14))%*%t(b))
round(fit.my$beta_hat,2)
round(fit.my$beta_hat_se,3)
is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))
diag(cor(fit.my$weight,fit.music$Weight.gene,use = "complete.obs"))
```

Scaling of y, selection of nu is tricky.

Use music X,weights,variance to calculate weights: 

```{r}
# use everything from music
# music weights
fit.my = wols(fit.music$X,datax$y,fit.music$Sigma,b=t(as.matrix(fit.music$Est.prop.weighted))*100,
              w.mode = 'res+bvb',adj.v = F,X_array = design.mat$X_array,
              hc.type = 'hc0',marker = rownames(fit.music$Weight.gene),nu=1e-4,scale.y = T)
rmse(fit.my$beta_hat,(rep(1,14))%*%t(b))
round(fit.my$beta_hat,2)
diag(cor(fit.my$weight,fit.music$Weight.gene,use = "complete.obs"))
```



other hc2, hc3

```{r}
fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc2',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = T)
is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))
round(fit.my$beta_hat_se,3)

fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc3',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = T)
is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))
round(fit.my$beta_hat_se,3)

```

do not use weights when calculating variance

```{r}
fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc0',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = F)
is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))
round(fit.my$beta_hat_se,3)

fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc2',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = F)
is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))
round(fit.my$beta_hat_se,3)

fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc3',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = F)
is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))
round(fit.my$beta_hat_se,3)

```

```{r,eval=F}
plot(fit.my$weight[,3])
quantile((fit.my$weight[,3]),na.rm=TRUE)
order(fit.my$weight[,3],decreasing = T)[1:10]

# top 10 weighted samples

round(fit.my$weight[order(fit.my$weight[,3],decreasing = T)[1:10],3],7)

round(fit.my$X[order(fit.my$weight[,3],decreasing = T)[1:10],],4)

```


**gene weights comparison with cibersort x**

```{r}
seger_sig_cibersort_output = read.delim("data/pancreas/seger_sig_cibersort_output.txt", quote="")
sig_genes = seger_sig_cibersort_output[,1]
length(sig_genes)


genes_all = rownames(fit.music$Weight.gene)
w_cibersort = rep(0,length(genes_all))
names(w_cibersort) = genes_all
w_cibersort[names(w_cibersort)%in%sig_genes] = 1


cor(fit.my$weight[,1],w_cibersort,use='pairwise.complete.obs')
cor(fit.music$Weight.gene[,1],w_cibersort,use='pairwise.complete.obs')

cor(fit.my$weight[,2],w_cibersort,use='pairwise.complete.obs')
cor(fit.music$Weight.gene[,2],w_cibersort,use='pairwise.complete.obs')

cor(fit.my$weight[,3],w_cibersort,use='pairwise.complete.obs')
cor(fit.music$Weight.gene[,3],w_cibersort,use='pairwise.complete.obs')

#plot(w_cibersort,fit.music$Weight.gene[,1])
#cibersort result

seger_cibersort_Results = read.delim("data/pancreas/seger_cibersort_Results.txt", quote="")
cibersort_est = seger_cibersort_Results[,2:5]
cibersort_est = cibersort_est[,c(2,4,1,3)]
round(cibersort_est,2)
rmse(as.matrix(cibersort_est),(rep(1,14))%*%t(b))

```


## Baron as reference

```{r}
# fit music
baron.es = sce_to_es(baron)
cell_size = data.frame(type = c('alpha', 'beta', 'delta', 'gamma'), size=c(1,1,1,1))
fit.music = music_prop(bulk_mat_to_es(bulk.xin$y),baron.es,clusters = 'cell_type', samples = 'individual',
                       select.ct = cell_types,cell_size=cell_size,verbose = FALSE,nu=1e-4)
rmse(fit.music$Est.prop.weighted,(rep(1,14))%*%t(b))
round(fit.music$Est.prop.weighted,2)

fit.music2 = music_prop(bulk_mat_to_es(bulk.xin$y),baron.es,clusters = 'cell_type', samples = 'individual',
                       select.ct = cell_types,cell_size=cell_size,verbose = FALSE,nu=1e-4,normalize = TRUE)
rmse(fit.music2$Est.prop.weighted,(rep(1,14))%*%t(b))
round(fit.music2$Est.prop.weighted,2)
```

```{r}

datax = set_data_decon(y = bulk.xin$y, Y = baron,
                       cell_types = cell_types, 
                       gene_thresh = 0,max_count_quantile = 1,w=1)

design.mat = scRef_multi_proc(datax$Y,datax$cell_type_idx,datax$indi_idx,
                                    estimator="separate",est_sigma2 = FALSE)

fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc0',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = T)
rmse(fit.my$beta_hat,(rep(1,14))%*%t(b))
round(fit.my$beta_hat,2)
round(fit.my$beta_hat_se,3)
is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))
```

```{r}
diag(cor(fit.my$weight,fit.music$Weight.gene,use = "complete.obs"))
diag(cor(fit.my$X,fit.music$X,use = "complete.obs"))
diag(cor(fit.my$Vg[,c(1,6,11,16)],fit.music$Sigma,use = "complete.obs"))
```



```{r}
# use everything from music
# music weights
fit.my = wols(fit.music$X,datax$y,fit.music$Sigma,b=t(as.matrix(fit.music$Est.prop.weighted))*100,
              w.mode = 'res+bvb',adj.v = F,X_array = design.mat$X_array,
              hc.type = 'hc0',marker = rownames(fit.music$Weight.gene),nu=1e-4,scale.y = T)
rmse(fit.my$beta_hat,(rep(1,14))%*%t(b))
round(fit.my$beta_hat,2)
diag(cor(fit.my$weight,fit.music$Weight.gene,use = "complete.obs"))
```



do not use weights when calculating variance

```{r}
fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc0',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = F)
round(fit.my$beta_hat_se,3)
is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))

fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc2',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = F)
round(fit.my$beta_hat_se,3)
is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))

fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc3',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = F)
round(fit.my$beta_hat_se,3)
is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,14))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))
```


**gene weights comparison with cibersort x**

```{r}
baron_sig_cibersort_output = read.delim("data/pancreas/baron_sig_cibersort_output.txt", quote="")
sig_genes = baron_sig_cibersort_output[,1]
length(sig_genes)


genes_all = rownames(fit.music$Weight.gene)
w_cibersort = rep(0,length(genes_all))
names(w_cibersort) = genes_all
w_cibersort[names(w_cibersort)%in%sig_genes] = 1


cor(fit.my$weight[,1],w_cibersort,use='complete.obs')
cor(fit.music$Weight.gene[,1],w_cibersort,use='complete.obs')

cor(fit.my$weight[,2],w_cibersort,use='complete.obs')
cor(fit.music$Weight.gene[,2],w_cibersort,use='complete.obs')

cor(fit.my$weight[,3],w_cibersort,use='complete.obs')
cor(fit.music$Weight.gene[,3],w_cibersort,use='complete.obs')

#plot(w_cibersort,fit.music$Weight.gene[,1])
#cibersort result

baron_cibersort_Results = read.delim("data/pancreas/baron_cibersort_Results.txt", quote="")
cibersort_est = baron_cibersort_Results[,2:5]
cibersort_est = cibersort_est[,c(3,1,2,4)]
round(cibersort_est,2)
rmse(as.matrix(cibersort_est),(rep(1,14))%*%t(b))

```


## Create bulk data(with technical noise) using Seger dataset.

```{r}
datax = set_data_decon(Y = seger,cell_types = cell_types, gene_thresh = 0,max_count_quantile = 1,w=1)
design.mat = scRef_multi_proc(datax$Y,datax$cell_type_idx,datax$indi_idx,estimator="separate",est_sigma2 = FALSE)
# create bulk data
b = c(0.1,0.1,0.3,0.5)
bulk.seger = createBulk_tech.noise(design.mat$X_array,b,gene_names = rownames(seger),seed=12345,ls_per_gene = 500)
```


## Xin as reference


```{r}
# fit music
xin.es = sce_to_es(xin_raw)
cell_size = data.frame(type = c('alpha', 'beta', 'delta', 'gamma'), size=c(1,1,1,1))
fit.music = music_prop(bulk_mat_to_es(bulk.seger$y),xin.es,clusters = 'cell_type', samples = 'individual',
                       select.ct = cell_types,cell_size=cell_size,verbose = FALSE,nu=1e-4,normalize = F)
rmse(fit.music$Est.prop.weighted,(rep(1,10))%*%t(b))
round(fit.music$Est.prop.weighted,2)

```


```{r}
datax = set_data_decon(y = bulk.seger$y, Y = xin_raw[,-rm.indi.idx],
                       cell_types = cell_types, 
                       gene_thresh = 0,max_count_quantile = 1,w=1,filter.gene = F)

design.mat = scRef_multi_proc(datax$Y,datax$cell_type_idx,datax$indi_idx,
                                    estimator="separate",est_sigma2 = FALSE)

fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc0',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = T)
rmse(fit.my$beta_hat,(rep(1,10))%*%t(b))
round(fit.my$beta_hat,2)
round(fit.my$beta_hat_se,3)
is.covered((rep(1,10))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,10))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))


fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc2',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = T)
rmse(fit.my$beta_hat,(rep(1,10))%*%t(b))
round(fit.my$beta_hat,2)
round(fit.my$beta_hat_se,3)
is.covered((rep(1,10))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,10))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))


fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc3',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = T)
rmse(fit.my$beta_hat,(rep(1,10))%*%t(b))
round(fit.my$beta_hat,2)
round(fit.my$beta_hat_se,3)
is.covered((rep(1,10))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,10))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))

```


```{r}
ii = which(rownames(fit.music$Weight.gene)%in%rownames(fit.my$weight))
diag(cor(fit.my$weight,fit.music$Weight.gene[ii,],use = "complete.obs"))
diag(cor(fit.my$X,fit.music$X[ii,],use = "pairwise.complete.obs"))
diag(cor(fit.my$Vg[,c(1,6,11,16)],fit.music$Sigma[ii,],use = "pairwise.complete.obs"))
```


```{r}
fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc0',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = F)
rmse(fit.my$beta_hat,(rep(1,10))%*%t(b))
round(fit.my$beta_hat,2)
round(fit.my$beta_hat_se,3)
is.covered((rep(1,10))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,10))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))

fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc2',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = F)
rmse(fit.my$beta_hat,(rep(1,10))%*%t(b))
round(fit.my$beta_hat,2)
round(fit.my$beta_hat_se,3)
is.covered((rep(1,10))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,10))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))

fit.my = wols(design.mat$X,datax$y,design.mat$Vg,b=NULL,
              w.mode = 'res+ref_var',adj.v = T,X_array = design.mat$X_array,
              hc.type = 'hc3',marker = rownames(fit.music$Weight.gene),
              nu=1e-4,scale.y = F,use.weight.for.var = F)
rmse(fit.my$beta_hat,(rep(1,10))%*%t(b))
round(fit.my$beta_hat,2)
round(fit.my$beta_hat_se,3)
is.covered((rep(1,10))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se)
mean(is.covered((rep(1,10))%*%t(b),fit.my$beta_hat,fit.my$beta_hat_se))
```







