---
title: "Understanding Least Squares"
author: "DongyueXie"
date: "2020-04-07"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

If $y\sim\text{Poisson}(X\beta)$ and $Z\sim\text{Poisson}(X)$, then OLS is biased and not consistent.

```{r}
b = 1:3
p=length(b)
nn = 100
n_list = round(seq(30,300,length.out = nn))
nreps=100
set.seed(12345)

mean_olsx = matrix(nrow=nn,ncol=p)
mean_olsz = matrix(nrow=nn,ncol=p)
mean_olsz_adj = matrix(nrow=nn,ncol=p)

var_olsx = matrix(nrow=nn,ncol=p)
var_olsz = matrix(nrow=nn,ncol=p)
var_olsz_adj = matrix(nrow=nn,ncol=p)

mse_olsx = matrix(nrow=nn,ncol=p)
mse_olsz = matrix(nrow=nn,ncol=p)
mse_olsz_adj = matrix(nrow=nn,ncol=p)

for(j in 1:length(n_list)){
  
  n = n_list[j]
  
  b_olsx = matrix(nrow=nreps,ncol=p)
  b_olsz = matrix(nrow=nreps,ncol=p)
  b_olsz_adj = matrix(nrow=nreps,ncol=p)
  
  for(i in 1:nreps){
    X = matrix(rpois(n*p,10),nrow=n)
    Z = matrix(rpois(n*p,X),ncol=p)
    y = rpois(n,X%*%b)
    
    b_olsx[i,] = solve(t(X)%*%X)%*%t(X)%*%y
    b_olsz[i,] = solve(t(Z)%*%Z)%*%t(Z)%*%y
    b_olsz_adj[i,] = solve(t(Z)%*%Z - diag(colSums(Z)))%*%t(Z)%*%y
  }
  
  mean_olsx[j,] = apply(b_olsx,2,mean)
  mean_olsz[j,] = apply(b_olsz,2,mean)
  mean_olsz_adj[j,] = apply(b_olsz_adj,2,mean)
  
  var_olsx[j,] = apply(b_olsx,2,var)
  var_olsz[j,] = apply(b_olsz,2,var)
  var_olsz_adj[j,] = apply(b_olsz_adj,2,var)
}

```

```{r,fig.width=9,fig.height=8}
par(mfrow=c(3,2))
for(i in 1:p){
  plot(n_list,mean_olsx[,i],xlab='n',ylab='mean',
       ylim=range(c(mean_olsx[,i],mean_olsz[,i])),main=paste('beta',i),type='l')
  lines(n_list,mean_olsz[,i],col=2)
  #lines(n_list,mean_olsz_adj[,i],col=4)
  abline(h=b[i],lty=2)
  
  plot(n_list,var_olsx[,i],xlab='n',ylab='var',
       ylim=range(c(0,var_olsx[,i],var_olsz[,i])),main=paste('beta',i),type='l')
  lines(n_list,var_olsz[,i],col=2)
  #lines(n_list,var_olsz_adj[,i],col=4)
  abline(h=0,lty=2)
  legend('topright',c('ols with x','ols with z'),lty=c(1,1),col=c(1,2))
}
```

The plot shows that 1. ols with X is unbiased and consistent; 2. ols with Z, without adjustment, is biased and inconsistent.

How about the adjusted one?

```{r,fig.width=9,fig.height=8}
par(mfrow=c(3,2))
for(i in 1:p){
  plot(n_list,mean_olsz_adj[,i],xlab='n',ylab='mean',main=paste('beta',i),type='l',ylim=c(b[i]-0.5,b[i]+0.5))
  abline(h=b[i],lty=2)
  
  plot(n_list,var_olsz_adj[,i],xlab='n',ylab='var',main=paste('beta',i),type='l',ylim=c(0,0.5))
  abline(h=0,lty=2)
}
```

It dose not perform well when sample size is small. Is there any small sample ajustment? It's unbiased but the variance is much larger(5 times of ols with z).  
