---
title: "multiple testing of correlations"
author: "Dongyue Xie"
date: "2021-05-15"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

We review the one sample case, section 4 in the paper [Cai and Liu(2016)](https://www.tandfonline.com/doi/abs/10.1080/01621459.2014.999157).

Suppose $n$ samples from p-dimensional distribution with mean $\mu$ and correlation matrix $R$ are observed and the goal is to simultaneously test $H_{0ij}:\rho_{ij} = 0$. It's quivalent to $H_{0ij}:\sigma_{ij} = 0$. A test statistics is 
\[T_{ij} = \frac{\sum_{k=1}^n (x_{ki}-\bar x_i)(x_{kj}-\bar x_j)}{\sqrt{n\hat\theta_{ij}}},\]
where $\bar x_i = \frac{1}{n}\sum_k x_{k,i}$ and 
\[\hat\theta_{ij} = \frac{1}{n}\sum_k ((x_{ki}-\bar x_i)(x_{kj}-\bar x_j)-\hat\sigma_{ij})^2\]

Define 
\[\hat t = \inf\{0\leq t\leq b_p:\frac{G(t)(p^2-p)/2}{\max\{\sum_{i<j} I(|T_{ij}\geq t|),1\}}\},\]
where $G(t) = 2(1-\Phi(t))$, $b_p = \sqrt{4\log p-2\log(\log p)}$. If $\hat t$ does not exist, then set $\hat t = \sqrt{4\log p}$.

Reject $H_{0ij}$ when $|T_{ij}|\geq \hat t$. 

## Implementation

The two quantities need calculating is the sample covariance matrix $S$ and $\sum_{k=1}^n[(x_k-\mu)(x_k-\mu)^T]^2$ where the square is taken elementwisely.

```{r}
library(Matrix)
library(Rfast)
#'@param X sample by condition matrix
#'@param 
multipletesting_correlation = function(X,alpha=0.05){
  n = nrow(X)
  p = ncol(X)
  
  X.center = scale(X,center=TRUE,scale=FALSE)
  S = cova(X.center,center = TRUE)
  
  # calc S2
  
  S2 = 0
  for(k in 1:n){
    S2 = S2+(tcrossprod(X.center[k,]))^2
  }
  
  # calc T statistics
  
  Tmat = S*(n-p)/sqrt(S2+(2*p-n)*S^2)
  
  # find t
  bp = sqrt(4*log(p)-2*log(log(p)))
  a = (p^2-p)/2
  Tvec = Tmat[lower.tri(Tmat)]
  t_vec = sort(abs(Tvec),decreasing = FALSE)
  nt = length(t_vec)
  for(t in 1:nt){
    temp = 2*(1-pnorm(t_vec[t]))*a/(nt-t+1)
    if(temp<=alpha){
      break
    }
  }
  if(t_vec[t]>bp){
    thresh = sqrt(4*log(p))
  }else{
    thresh = t_vec[t]
  }
  list(S=S,
       S2=S2,
       Tmat=Tmat,
       Tvec = Tvec,
       thresh=thresh,
       alpha=alpha)
}


```


Example, see the null distribution of test statistics

```{r}
n = 3000
p = 100
X = matrix(rnorm(n*p),nrow=n,ncol=p)
out = multipletesting_correlation(X)
#which(abs(out$Tvec)>=out$thresh)
h=hist(out$Tvec,breaks = 100)
xfit<-seq(min(out$Tvec),max(out$Tvec),length=100)
yfit<-dnorm(xfit,mean=0,sd=1)
yfit <- yfit*diff(h$mids[1:2])*length(out$Tvec)
lines(xfit, yfit, col="blue", lwd=2)
```


```{r}
n = 300
p = 100
X = matrix(rnorm(n*p),nrow=n,ncol=p)
out = multipletesting_correlation(X)
#which(abs(out$Tvec)>=out$thresh)
h=hist(out$Tvec,breaks = 100)
xfit<-seq(min(out$Tvec),max(out$Tvec),length=100)
yfit<-dnorm(xfit,mean=0,sd=1)
yfit <- yfit*diff(h$mids[1:2])*length(out$Tvec)
lines(xfit, yfit, col="blue", lwd=2)
```

```{r}
n = 30
p = 100
X = matrix(rnorm(n*p),nrow=n,ncol=p)
out = multipletesting_correlation(X)
#which(abs(out$Tvec)>=out$thresh)
h=hist(out$Tvec,breaks = 100)
xfit<-seq(min(out$Tvec),max(out$Tvec),length=100)
yfit<-dnorm(xfit,mean=0,sd=1)
yfit <- yfit*diff(h$mids[1:2])*length(out$Tvec)
lines(xfit, yfit, col="blue", lwd=2)
```






```{r}
n = 300
p = 1000
X = matrix(rnorm(n*p),nrow=n,ncol=p)
out = multipletesting_correlation(X)
#which(abs(out$Tvec)>=out$thresh)
h=hist(out$Tvec,breaks = 100)
xfit<-seq(min(out$Tvec),max(out$Tvec),length=100)
yfit<-dnorm(xfit,mean=0,sd=1)
yfit <- yfit*diff(h$mids[1:2])*length(out$Tvec)
lines(xfit, yfit, col="blue", lwd=2)
```


```{r}
n = 300
p = 5000
X = matrix(rnorm(n*p),nrow=n,ncol=p)
out = multipletesting_correlation(X)
#which(abs(out$Tvec)>=out$thresh)
h=hist(out$Tvec,breaks = 100)
xfit<-seq(min(out$Tvec),max(out$Tvec),length=100)
yfit<-dnorm(xfit,mean=0,sd=1)
yfit <- yfit*diff(h$mids[1:2])*length(out$Tvec)
lines(xfit, yfit, col="blue", lwd=2)
```


