---
title: "Investigate multiple individuals estimation"
author: "DongyueXie"
date: "2020-05-13"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

In [previous analysis](multiIndSimulation.html), it seems that the adjustment did not lead to better coverage. 

Let's set $\sigma^2$ all equal and see how well $\sigma^2$ is estiamted.

If set sigma2 = 1/G, and generate individual gene relative expresison from gamma distribution, then a lot of relative expressions are very close to 0. So even the bulk data vector y is very sparse.


## fix tau2 = 1/G^2 and sigma2 = 1/G^2

### both are known

```{r}
source('code/deconference.R')
source('code/simu_func_multi.R')
```

```{r}

library(gtools)
set.seed(12345)

G = 500
K = 4
b = 1:K
b = b/sum(b)

ref = t(rdirichlet(K,rep(1,G)))
bulk_lib_size = 50
n_indi = 10
nk=50
sc_lib_size = 0.2

Ng = G
sigma2 = 1/Ng^2
tau2 = 1/Ng^2

results1 = simu_study_multiInd(ref,Ng,(b)/sum(b),
                                     n_indi = n_indi,
                                     sigma2known=T,
                                     tau2known = T,
                                     x_estimator = 'separate',
                                     tau2=tau2,
                                     sigma2=sigma2,
                                     addw=TRUE,
                                     bulk_lib_size=bulk_lib_size,
                                     sc_lib_size=sc_lib_size,
                                     meta_var = 'adjust',
                                     est_sigma2 = F,
                                     eps=1e-8,
                                     nk=nk,nreps=50)
```


```{r}

results1$covergae_adj
results1$covergae_unadj

results1$mean_est_adj
results1$mean_est_unadj

results1$se_est_adj
apply(results1$se_adj,2,mean)

results1$se_est_unadj
apply(results1$se_unadj,2,mean)
```

```{r}
mean(results1$data_sparsity)
```

### tau2 is unknown

```{r}
set.seed(12345)

G = 500
K = 4
b = 1:K
b = b/sum(b)

ref = t(rdirichlet(K,rep(1,G)))
bulk_lib_size = 50
n_indi = 10
nk=50
sc_lib_size = 0.2

Ng = G
sigma2 = 1/Ng^2
tau2 = 1/Ng^2

results1 = simu_study_multiInd(ref,Ng,(b)/sum(b),
                                     n_indi = n_indi,
                                     sigma2known=T,
                                     tau2known = F,
                                     x_estimator = 'separate',
                                     tau2=tau2,
                                     sigma2=sigma2,
                                     addw=TRUE,
                                     bulk_lib_size=bulk_lib_size,
                                     sc_lib_size=sc_lib_size,
                                     meta_var = 'adjust',
                                     est_sigma2 = F,
                                     eps=1e-8,
                                     nk=nk,nreps=50)
```


```{r}

results1$covergae_adj
results1$covergae_unadj

results1$mean_est_adj
results1$mean_est_unadj

results1$se_est_adj
apply(results1$se_adj,2,mean)

results1$se_est_unadj
apply(results1$se_unadj,2,mean)
```

### sigma2 and tau2 are unknown  and use meta analysis

```{r}
set.seed(12345)

G = 500
K = 4
b = 1:K
b = b/sum(b)

ref = t(rdirichlet(K,rep(1,G)))
bulk_lib_size = 50
n_indi = 10
nk=50
sc_lib_size = 0.2

Ng = G
sigma2 = 1/Ng^2
tau2 = 1/Ng^2

results1 = simu_study_multiInd(ref,Ng,(b)/sum(b),
                                     n_indi = n_indi,
                                     sigma2known=F,
                                     tau2known = F,
                                     x_estimator = 'separate',
                                     tau2=tau2,
                                     sigma2=sigma2,
                                     addw=TRUE,
                                     bulk_lib_size=bulk_lib_size,
                                     sc_lib_size=sc_lib_size,
                                     meta_var = 'adjust',
                                     est_sigma2 = T,
                                     eps=1e-8,
                                     nk=nk,nreps=50)
```


```{r}

results1$covergae_adj
results1$covergae_unadj

results1$mean_est_adj
results1$mean_est_unadj

results1$se_est_adj
apply(results1$se_adj,2,mean)

results1$se_est_unadj
apply(results1$se_unadj,2,mean)
```


how well are the estimate of sigma2?

```{r}
sigma2
mean(results1$sigma2_hat[[1]])
mean(results1$sigma2_hat[[2]])
```

### sigma2 and tau2 are unknown and directly use sample cov among individuals

```{r}
set.seed(12345)

G = 500
K = 4
b = 1:K
b = b/sum(b)

ref = t(rdirichlet(K,rep(1,G)))
bulk_lib_size = 50
n_indi = 10
nk=50
sc_lib_size = 0.2

Ng = G
sigma2 = 1/Ng^2
tau2 = 1/Ng^2

results1 = simu_study_multiInd(ref,Ng,(b)/sum(b),
                                     n_indi = n_indi,
                                     sigma2known=F,
                                     tau2known = F,
                                     x_estimator = 'separate',
                                     tau2=tau2,
                                     sigma2=sigma2,
                                     addw=TRUE,
                                     bulk_lib_size=bulk_lib_size,
                                     sc_lib_size=sc_lib_size,
                                     meta_var = 'adjust',
                                     est_sigma2 = F,
                                     eps=1e-8,
                                     nk=nk,nreps=50)
```


```{r}

results1$covergae_adj
results1$covergae_unadj

results1$mean_est_adj
results1$mean_est_unadj

results1$se_est_adj
apply(results1$se_adj,2,mean)

results1$se_est_unadj
apply(results1$se_unadj,2,mean)
```


## fix tau2 = 1/G and sigma2 = 1/G^2

### both are known

Now increase variance among cells within individual

```{r}

library(gtools)
set.seed(12345)

G = 500
K = 4
b = 1:K
b = b/sum(b)

ref = t(rdirichlet(K,rep(1,G)))
bulk_lib_size = 50
n_indi = 10
nk=50
sc_lib_size = 0.2

Ng = G
sigma2 = 1/Ng^2
tau2 = 1/Ng

results1 = simu_study_multiInd(ref,Ng,(b)/sum(b),
                                     n_indi = n_indi,
                                     sigma2known=T,
                                     tau2known = T,
                                     x_estimator = 'separate',
                                     tau2=tau2,
                                     sigma2=sigma2,
                                     addw=TRUE,
                                     bulk_lib_size=bulk_lib_size,
                                     sc_lib_size=sc_lib_size,
                                     meta_var = 'adjust',
                                     est_sigma2 = T,
                                     eps=1e-8,
                                     nk=nk,nreps=50)
```


```{r}

results1$covergae_adj
results1$covergae_unadj

results1$mean_est_adj
results1$mean_est_unadj

results1$se_est_adj
apply(results1$se_adj,2,mean)

results1$se_est_unadj
apply(results1$se_unadj,2,mean)
```

```{r}
mean(results1$data_sparsity)
```


Data is very sparse. 98 percent of Y are 0. The problem is that :

When a gene only appears in one cell type in one individual, then it's estimated relative expression would be mostly 0 and we specifying variance to known so obviously varaince are too large for adjustment.


### both are unknown and use meta analysis

```{r}

library(gtools)
set.seed(12345)

G = 500
K = 4
b = 1:K
b = b/sum(b)

ref = t(rdirichlet(K,rep(1,G)))
bulk_lib_size = 50
n_indi = 10
nk=50
sc_lib_size = 0.2

Ng = G
sigma2 = 1/Ng^2
tau2 = 1/Ng

results1 = simu_study_multiInd(ref,Ng,(b)/sum(b),
                                     n_indi = n_indi,
                                     sigma2known=F,
                                     tau2known = F,
                                     x_estimator = 'separate',
                                     tau2=tau2,
                                     sigma2=sigma2,
                                     addw=TRUE,
                                     bulk_lib_size=bulk_lib_size,
                                     sc_lib_size=sc_lib_size,
                                     meta_var = 'adjust',
                                     est_sigma2 = T,
                                     eps=1e-8,
                                     nk=nk,nreps=50)
```


```{r}

results1$covergae_adj
results1$covergae_unadj

results1$mean_est_adj
results1$mean_est_unadj

results1$se_est_adj
apply(results1$se_adj,2,mean)

results1$se_est_unadj
apply(results1$se_unadj,2,mean)
```

how well are the estimate of sigma2?

```{r}
sigma2
mean(results1$sigma2_hat[[1]])
mean(results1$sigma2_hat[[2]])
```


### sigma2 and tau2 are unknown and directly use sample cov among individuals

```{r}
set.seed(12345)

G = 500
K = 4
b = 1:K
b = b/sum(b)

ref = t(rdirichlet(K,rep(1,G)))
bulk_lib_size = 50
n_indi = 10
nk=50
sc_lib_size = 0.2

Ng = G
sigma2 = 1/Ng^2
tau2 = 1/Ng

results1 = simu_study_multiInd(ref,Ng,(b)/sum(b),
                                     n_indi = n_indi,
                                     sigma2known=F,
                                     tau2known = F,
                                     x_estimator = 'separate',
                                     tau2=tau2,
                                     sigma2=sigma2,
                                     addw=TRUE,
                                     bulk_lib_size=bulk_lib_size,
                                     sc_lib_size=sc_lib_size,
                                     meta_var = 'adjust',
                                     est_sigma2 = F,
                                     eps=1e-8,
                                     nk=nk,nreps=50)
```


```{r}

results1$covergae_adj
results1$covergae_unadj

results1$mean_est_adj
results1$mean_est_unadj

results1$se_est_adj
apply(results1$se_adj,2,mean)

results1$se_est_unadj
apply(results1$se_unadj,2,mean)
```

Chnage to aggregate estimator?

```{r}
set.seed(12345)

G = 500
K = 4
b = 1:K
b = b/sum(b)

ref = t(rdirichlet(K,rep(1,G)))
bulk_lib_size = 50
n_indi = 10
nk=50
sc_lib_size = 0.2

Ng = G
sigma2 = 1/Ng^2
tau2 = 1/Ng

results1 = simu_study_multiInd(ref,Ng,(b)/sum(b),
                                     n_indi = n_indi,
                                     sigma2known=F,
                                     tau2known = F,
                                     x_estimator = 'aggregate',
                                     tau2=tau2,
                                     sigma2=sigma2,
                                     addw=TRUE,
                                     bulk_lib_size=bulk_lib_size,
                                     sc_lib_size=sc_lib_size,
                                     meta_var = 'adjust',
                                     est_sigma2 = F,
                                     eps=1e-8,
                                     nk=nk,nreps=50)
```


```{r}

results1$covergae_adj
results1$covergae_unadj

results1$mean_est_adj
results1$mean_est_unadj

results1$se_est_adj
apply(results1$se_adj,2,mean)

results1$se_est_unadj
apply(results1$se_unadj,2,mean)
```
