---
title: "multiple testing of correlations"
author: "Dongyue Xie"
date: "2021-05-15"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

We review the one sample case, section 4 in the paper [Cai and Liu(2016)](https://www.tandfonline.com/doi/abs/10.1080/01621459.2014.999157).

Suppose $n$ samples from p-dimensional distribution with mean $\mu$ and correlation matrix $R$ are observed and the goal is to simultaneously test $H_{0ij}:\rho_{ij} = 0$. It's quivalent to $H_{0ij}:\sigma_{ij} = 0$. A test statistics is 
\[T_{ij} = \frac{\sum_{k=1}^n (x_{ki}-\bar x_i)(x_{kj}-\bar x_j)}{\sqrt{n\hat\theta_{ij}}},\]
where $\bar x_i = \frac{1}{n}\sum_k x_{k,i}$ and 
\[\hat\theta_{ij} = \frac{1}{n}\sum_k ((x_{ki}-\bar x_i)(x_{kj}-\bar x_j)-\hat\sigma_{ij})^2\]

Define 
\[\hat t = \inf\{0\leq t\leq b_p:\frac{G(t)(p^2-p)/2}{\max\{\sum_{i<j} I(|T_{ij}\geq t|),1\}}\},\]
where $G(t) = 2(1-\Phi(t))$, $b_p = \sqrt{4\log p-2\log(\log p)}$. If $\hat t$ does not exist, then set $\hat t = \sqrt{4\log p}$.

Reject $H_{0ij}$ when $|T_{ij}|\geq \hat t$. 

```{r}
library(Matrix)
library(Rfast)
#'@param X sample by condition matrix
#'@param 
multipletesting_correlation = function(X,alpha=0.05){
  n = nrow(X)
  p = ncol(X)
  S = cova(X)
  mean.col = colmeans(X)
  Theta = Matrix(0,nrow=p,ncol=p,sparse = TRUE)
  for(i in 1:(p-1)){
    for(j in (1+i):p){
      Theta[i,j] = mean(((X[,i]-mean.col[i])*(X[,j]-mean.col[j]) - S[i,j])^2)
    }
  }
  Tmat = Matrix(0,nrow=p,ncol=p,sparse = TRUE)
  for(i in 1:(p-1)){
    for(j in (1+i):p){
      Tmat[i,j] = sum((X[,i]-mean.col[i])*(X[,j]-mean.col[j]))/sqrt(n*Theta[i,j])
    }
  }
  # find t
  bp = sqrt(4*log(p)-2*log(log(p)))
  a = (p^2-p)/2
  Tvec = Tmat[(Tmat!=0)]
  t_vec = sort(abs(Tvec),decreasing = FALSE)
  nt = length(t_vec)
  for(t in 1:nt){
    temp = 2*(1-pnorm(t_vec[t]))*a/(nt-t+1)
    if(temp<=alpha){
      break
    }
  }
  if(t_vec[t]>bp){
    thresh = sqrt(4*log(p))
  }else{
    thresh = t_vec[t]
  }
  list(S=S,
       Tmat=Tmat,
       Tvec = Tvec,
       t=thresh,
       alpha=alpha)
}


```


Example

```{r}
n = 30
p = 100
X = MASS::mvrnorm(n,rep(0,p),diag(p))
out = multipletesting_correlation(X)
which(abs(out$Tvec)>=out$t)

X = MASS::mvrnorm(n,rep(0,p),diag(p))
out = multipletesting_correlation(X)
which(abs(out$Tvec)>=out$t)
```


