---
title: "music_weights"
author: "DongyueXie"
date: "2020-09-12"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction



```{r,warning=F,message=F}
library(MuSiC)
library(Biobase)
library(xbioc)
library(SingleCellExperiment)
library(scater)
library(GEOquery)
```

```{r}
seger <- readRDS("~/deconference/data/pancreas/segerstolpe_raw.rds")
enge = readRDS("data/pancreas/enge_raw.rds")

create_sce_from_counts = function(counts, colData, rowData = NULL) {
  if(is.null(rowData)) {
    sceset <- SingleCellExperiment(assays = list(counts = as.matrix(counts)),
                                   colData = colData)
  } else {
    sceset <- SingleCellExperiment(assays = list(counts = as.matrix(counts)),
                                   colData = colData,
                                   rowData = rowData)
  }
  # use gene names as feature symbols
  rowData(sceset)$feature_symbol <- rownames(sceset)
  # remove features with duplicated names
  if(is.null(rowData)) {
    sceset <- sceset[!duplicated(rowData(sceset)$feature_symbol), ]
  }
  # QC
  isSpike(sceset, "ERCC") <- grepl("^ERCC-", rownames(sceset))
  sceset <- calculateQCMetrics(sceset, feature_controls = list("ERCC" = isSpike(sceset, "ERCC")))
  return(sceset)
}

create_bulk = function(data.obj,indis,cell_types,min_prop=0.15){
  G = nrow(data.obj)
  bulk_counts = matrix(nrow=G,ncol=length(indis))
  col_condition = c()
  bulk_prop = c()
  bulk_ncell = c()
  #bulk_prop = matrix(nrow=length(indis),ncol = length(cell_types))
  for(i in 1:length(indis)){
    idx.temp = which(data.obj$individual==indis[i]&data.obj$cell_type%in%cell_types)
    cell_counts = table(data.obj$cell_type[idx.temp])
    
    if(!is.null(min_prop)){
      
      min_count = min(cell_counts)
    max_cell = floor(min(cell_counts)/min_prop)
    if(max_cell<sum(cell_counts)){
      
      to_rm = sum(cell_counts) - max_cell
      for(j in 1:length(cell_counts)){
        rm.celltype = names(cell_counts)[which.max(cell_counts)]
        candidate.rm.idx = which(data.obj[,idx.temp]$cell_type==rm.celltype)
        if(to_rm>(max(cell_counts)-min_count)){
          idx.temp = idx.temp[-candidate.rm.idx[sample(1:length(candidate.rm.idx),max(cell_counts)-min_count)]]
          to_rm = to_rm - max(cell_counts) + min_count
          cell_counts = table(data.obj$cell_type[idx.temp])
        }else{
          idx.temp = idx.temp[-candidate.rm.idx[sample(1:length(candidate.rm.idx),to_rm)]]
          cell_counts = table(data.obj$cell_type[idx.temp])
          break
        }
        
      }
      # while (to_rm>0&max(cell_counts)>(max_cell*min_prop)) {
      #   rm.celltype = names(cell_counts)[which.max(cell_counts)]
      #   candidate.rm.idx = which(data.obj[,idx.temp]$cell_type==rm.celltype)
      #   idx.temp = idx.temp[-candidate.rm.idx[sample(1:length(candidate.rm.idx),1)]]
      #   cell_counts = table(data.obj$cell_type[idx.temp])
      #   to_rm = to_rm - 1
      # }
    }
    
    }
    
    
    cell_prop = cell_counts/sum(cell_counts)
    bulk_prop = rbind(bulk_prop,cell_prop)
    bulk_ncell = rbind(bulk_ncell,cell_counts)
    bulk_counts[,i] = rowSums(counts(data.obj)[,idx.temp])
    #col_condition[i] = (data.obj$disease[which(data.obj$individual==indis[i])])[1]
    # bulk_prop[i,] = 
  }
  colnames(bulk_counts) = indis
  rownames(bulk_counts) = rownames(data.obj)
  rownames(bulk_ncell) = indis
  rownames(bulk_prop) = indis
  #ii = which(data.obj$individual%in%indis&data.obj$cell_type%in%cell_types)
  #bulk_prop = table(data.obj$individual[ii],data.obj$cell_type[ii])
  #bulk_prop = bulk_prop/rowSums(bulk_prop)
  coldata = DataFrame(individual = indis)
  out = create_sce_from_counts(bulk_counts,coldata)
  return(list(bulk = out,bulk_prop = bulk_prop,bulk_ncell=bulk_ncell))
}

summary_real = function(out,true_b){
  
  # compare fitted vs true
  b_hat = t(out$beta_hat)
  rownames(b_hat) = rownames(true_b)
  colnames(b_hat) = colnames(true_b)
  print(knitr::kable(
  list(t0 = round(b_hat,2), t1 = round(true_b,2)),
  caption = 'estimated vs true beta',
  booktabs = TRUE, valign = 't'))
  
  
  ci_length = t(round(out$beta_se,2))
  colnames(ci_length) = colnames(true_b)
  rownames(ci_length) = rownames(true_b)
  print(knitr::kable(ci_length,caption = 'beta hat se'))
  
  # summary statistics
  rmse = sqrt(mean((t(out$beta_hat)-true_b)^2))
  mad = mean(abs(t(out$beta_hat)-true_b))
  r2 = cor(c(t(out$beta_hat)),c(true_b))
  ss = rbind(c(rmse,mad,r2))
  colnames(ss) = c('rmse','mad','r2')
  rownames(ss) = NULL
  
  print(knitr::kable(ss,caption='error measure'))
  
  # coverage, length of CI
  ci = build_ci(out)
  coverage = c(true_b>=t(ci$ci_l))&(true_b<=t(ci$ci_r))
  
  print(knitr::kable(coverage,caption='coverage'))

  # two group test
  
  # groups = c(rep(1,3),rep(2,6))
  # g1 = which(groups==1)
  # g2 = which(groups==2)
  # test2 = two_group_test(out,groups)
  # 
  # test_out = rbind(round(colMeans(true_b[g1,]) - colMeans(true_b[g2,]),3),
  #                  round(test2$diff_group,3),
  #                  round(sqrt(diag(test2$V_tilde)),3),
  #                  round(test2$p_value,3))
  # rownames(test_out) = c('true diff','estimated diff','se','p value')
  # print(knitr::kable(test_out,caption= ' two group test'))
  
}

music_prop = function(bulk.eset, sc.eset, markers = NULL, clusters, samples, 
    select.ct = NULL, cell_size = NULL, ct.cov = FALSE, verbose = TRUE, 
    iter.max = 1000, nu = 1e-04, eps = 0.01, centered = FALSE, 
    normalize = FALSE, ...) 
{
    bulk.gene = rownames(bulk.eset)[rowMeans(exprs(bulk.eset)) != 
        0]
    bulk.eset = bulk.eset[bulk.gene, , drop = FALSE]
    if (is.null(markers)) {
        sc.markers = bulk.gene
    }
    else {
        sc.markers = intersect(bulk.gene, unlist(markers))
    }
    sc.basis = music_basis(sc.eset, non.zero = TRUE, markers = sc.markers, 
        clusters = clusters, samples = samples, select.ct = select.ct, 
        cell_size = cell_size, ct.cov = ct.cov, verbose = verbose)
    cm.gene = intersect(rownames(sc.basis$Disgn.mtx), bulk.gene)
    if (is.null(markers)) {
        if (length(cm.gene) < 0.2 * min(length(bulk.gene), nrow(sc.eset))) 
            stop("Too few common genes!")
    }
    else {
        if (length(cm.gene) < 0.2 * length(unlist(markers))) 
            stop("Too few common genes!")
    }
    if (verbose) {
        message(paste("Used", length(cm.gene), "common genes..."))
    }
    m.sc = match(cm.gene, rownames(sc.basis$Disgn.mtx))
    m.bulk = match(cm.gene, bulk.gene)
    D1 = sc.basis$Disgn.mtx[m.sc, ]
    M.S = colMeans(sc.basis$S, na.rm = T)
    if (!is.null(cell_size)) {
        if (!is.data.frame(cell_size)) {
            stop("cell_size paramter should be a data.frame with 1st column for cell type names and 2nd column for cell sizes")
        }
        else if (sum(names(M.S) %in% cell_size[, 1]) != length(names(M.S))) {
            stop("Cell type names in cell_size must match clusters")
        }
        else if (any(is.na(as.numeric(cell_size[, 2])))) {
            stop("Cell sizes should all be numeric")
        }
        my_ms_names <- names(M.S)
        cell_size <- cell_size[my_ms_names %in% cell_size[, 1], 
            ]
        M.S <- cell_size[match(my_ms_names, cell_size[, 1]), 
            ]
        M.S <- M.S[, 2]
        names(M.S) <- my_ms_names
    }
    Yjg = relative.ab(exprs(bulk.eset)[m.bulk, ])
    N.bulk = ncol(bulk.eset)
    if (ct.cov) {
        Sigma.ct = sc.basis$Sigma.ct[, m.sc]
        if (sum(Yjg[, i] == 0) > 0) {
            D1.temp = D1[Yjg[, i] != 0, ]
            Yjg.temp = Yjg[Yjg[, i] != 0, i]
            Sigma.ct.temp = Sigma.ct[, Yjg[, i] != 0]
            if (verbose) 
                message(paste(colnames(Yjg)[i], "has common genes", 
                  sum(Yjg[, i] != 0), "..."))
        }
        else {
            D1.temp = D1
            Yjg.temp = Yjg[, i]
            Sigma.ct.temp = Sigma.ct
            if (verbose) 
                message(paste(colnames(Yjg)[i], "has common genes", 
                  sum(Yjg[, i] != 0), "..."))
        }
        lm.D1.weighted = music.iter.ct(Yjg.temp, D1.temp, M.S, 
            Sigma.ct.temp, iter.max = iter.max, nu = nu, eps = eps, 
            centered = centered, normalize = normalize)
        Est.prop.allgene = rbind(Est.prop.allgene, lm.D1.weighted$p.nnls)
        Est.prop.weighted = rbind(Est.prop.weighted, lm.D1.weighted$p.weight)
        weight.gene.temp = rep(NA, nrow(Yjg))
        weight.gene.temp[Yjg[, i] != 0] = lm.D1.weighted$weight.gene
        Weight.gene = cbind(Weight.gene, weight.gene.temp)
        r.squared.full = c(r.squared.full, lm.D1.weighted$R.squared)
        Var.prop = rbind(Var.prop, lm.D1.weighted$var.p)
    }
    else {
        Sigma = sc.basis$Sigma[m.sc, ]
        valid.ct = (colSums(is.na(Sigma)) == 0) & (colSums(is.na(D1)) == 
            0) & (!is.na(M.S))
        if (sum(valid.ct) <= 1) {
            stop("Not enough valid cell type!")
        }
        if (verbose) {
            message(paste("Used", sum(valid.ct), "cell types in deconvolution..."))
        }
        D1 = D1[, valid.ct]
        M.S = M.S[valid.ct]
        D1 = D1/M.S[1]
        M.S = M.S/M.S[1]
        
        Sigma = Sigma[, valid.ct]
        Est.prop.allgene = NULL
        Est.prop.weighted = NULL
        Weight.gene = NULL
        r.squared.full = NULL
        Var.prop = NULL
        for (i in 1:N.bulk) {
            if (sum(Yjg[, i] == 0) > 0) {
                D1.temp = D1[Yjg[, i] != 0, ]
                Yjg.temp = Yjg[Yjg[, i] != 0, i]
                Sigma.temp = Sigma[Yjg[, i] != 0, ]
                if (verbose) 
                  message(paste(colnames(Yjg)[i], "has common genes", 
                    sum(Yjg[, i] != 0), "..."))
            }
            else {
                D1.temp = D1
                Yjg.temp = Yjg[, i]
                Sigma.temp = Sigma
                if (verbose) 
                  message(paste(colnames(Yjg)[i], "has common genes", 
                    sum(Yjg[, i] != 0), "..."))
            }
            lm.D1.weighted = music.iter(Yjg.temp, D1.temp, M.S, 
                Sigma.temp, iter.max = iter.max, nu = nu, eps = eps, 
                centered = centered, normalize = normalize)
            Est.prop.allgene = rbind(Est.prop.allgene, lm.D1.weighted$p.nnls)
            Est.prop.weighted = rbind(Est.prop.weighted, lm.D1.weighted$p.weight)
            weight.gene.temp = rep(NA, nrow(Yjg))
            weight.gene.temp[Yjg[, i] != 0] = lm.D1.weighted$weight.gene
            Weight.gene = cbind(Weight.gene, weight.gene.temp)
            r.squared.full = c(r.squared.full, lm.D1.weighted$R.squared)
            Var.prop = rbind(Var.prop, lm.D1.weighted$var.p)
        }
    }
    colnames(Est.prop.weighted) = colnames(D1)
    rownames(Est.prop.weighted) = colnames(Yjg)
    colnames(Est.prop.allgene) = colnames(D1)
    rownames(Est.prop.allgene) = colnames(Yjg)
    names(r.squared.full) = colnames(Yjg)
    colnames(Weight.gene) = colnames(Yjg)
    rownames(Weight.gene) = cm.gene
    colnames(Var.prop) = colnames(D1)
    rownames(Var.prop) = colnames(Yjg)
    return(list(Est.prop.weighted = Est.prop.weighted, Est.prop.allgene = Est.prop.allgene, 
        Weight.gene = Weight.gene, r.squared.full = r.squared.full, 
        Var.prop = Var.prop,X=D1,S=M.S))
}
```


## Xin as bulk, Seger as reference

run music

randomly select 3000 common genes, mainly for speed purpose

```{r}
XinT2D.eset <- readRDS("data/MuSiC/XinT2Deset.rds")
XinT2D.construct.full = bulk_construct(XinT2D.eset, clusters = 'cellType', samples = 'SubjectName')
XinT2D.construct.full$prop.real = relative.ab(XinT2D.construct.full$num.real, by.col = FALSE)
pheno_Data = data.frame(cellType = seger$cell_type,SubjectName = seger$individual)
rownames(pheno_Data) = colnames(seger)
seger.ref = ExpressionSet(assayData = counts(seger),phenoData = AnnotatedDataFrame(pheno_Data))

round(XinT2D.construct.full$prop.real,2)

set.seed(12)
cm.gene = intersect(rownames(seger.ref),rownames(XinT2D.construct.full$Bulk.counts))
cm.gene = cm.gene[sample(1:length(cm.gene),3000)]
g.bulk = match(cm.gene,rownames(XinT2D.construct.full$Bulk.counts))
g.ref = match(cm.gene,rownames(seger.ref))
Est.prop.xin = music_prop(XinT2D.construct.full$Bulk.counts[g.bulk,],seger.ref[g.ref,],clusters = 'cellType', samples = 'SubjectName', 
                          select.ct =  c('alpha', 'beta', 'delta', 'gamma'))
Eval_multi(prop.real = data.matrix(XinT2D.construct.full$prop.real), 
           prop.est = list(data.matrix(Est.prop.xin$Est.prop.weighted), 
                           data.matrix(Est.prop.xin$Est.prop.allgene)), 
           method.name = c('MuSiC', 'NNLS'))

round(Est.prop.xin$Est.prop.weighted,2)

set.seed(2)
cm.gene = intersect(rownames(seger.ref),rownames(XinT2D.construct.full$Bulk.counts))
cm.gene = cm.gene[sample(1:length(cm.gene),3000)]
g.bulk = match(cm.gene,rownames(XinT2D.construct.full$Bulk.counts))
g.ref = match(cm.gene,rownames(seger.ref))
Est.prop.xin = music_prop(XinT2D.construct.full$Bulk.counts[g.bulk,],seger.ref[g.ref,],clusters = 'cellType', samples = 'SubjectName', 
                          select.ct =  c('alpha', 'beta', 'delta', 'gamma'))
Eval_multi(prop.real = data.matrix(XinT2D.construct.full$prop.real), 
           prop.est = list(data.matrix(Est.prop.xin$Est.prop.weighted), 
                           data.matrix(Est.prop.xin$Est.prop.allgene)), 
           method.name = c('MuSiC', 'NNLS'))
rmse = function(x,y){sqrt(mean((x-y)^2))}

round(Est.prop.xin$Est.prop.weighted,2)
```

run deconference

weight = 1

```{r, results='asis'}
source('code/deconference_main.R')
datax = set_data_decon(exprs(XinT2D.construct.full$Bulk.counts),seger,marker_gene = rownames(Est.prop.xin$Weight.gene),cell_types = c('alpha', 'beta', 'delta', 'gamma'), gene_thresh = 0,max_count_quantile = 1,w=1)
out1 = deconference(datax,est_pop_var = TRUE,meta_mode = 'smooth',correction = T,verbose = TRUE,hc.type = 'hc3')
summary_real(out1,as.matrix(XinT2D.construct.full$prop.real))

out2 = deconference(datax,est_pop_var = TRUE,meta_mode = 'local',correction = T,verbose = TRUE,hc.type = 'hc3')
summary_real(out2,as.matrix(XinT2D.construct.full$prop.real))
```

compare estimated cell size

```{r}
round(Est.prop.xin$S,2)
round(out1$input$S,2)
```

now try use weights from music. note that music uses different weights for each individual in bulk sample, so we take the mean of all weights for each gene. 


```{r, results='asis'}
w = Est.prop.xin$Weight.gene
w = apply(w,2,function(z){z/sum(z,na.rm = T)})
w = rowMeans(w,na.rm = T)
datax = set_data_decon(exprs(XinT2D.construct.full$Bulk.counts),seger,marker_gene = rownames(Est.prop.xin$Weight.gene),cell_types = c('alpha', 'beta', 'delta', 'gamma'), gene_thresh = 0,max_count_quantile = 1,w=w)
out3 = deconference(datax,est_pop_var = TRUE,meta_mode = 'smooth',correction = T,verbose = TRUE,hc.type = 'hc3')
summary_real(out3,as.matrix(XinT2D.construct.full$prop.real))

out4 = deconference(datax,est_pop_var = TRUE,meta_mode = 'local',correction = T,verbose = TRUE,hc.type = 'hc3')
summary_real(out4,as.matrix(XinT2D.construct.full$prop.real))
```


```{r}
Eval_multi(prop.real = data.matrix(XinT2D.construct.full$prop.real), 
           prop.est = list(data.matrix(Est.prop.xin$Est.prop.weighted), 
                           data.matrix(Est.prop.xin$Est.prop.allgene),
                           t(out1$beta_hat),
                           t(out2$beta_hat),
                           t(out3$beta_hat),
                           t(out4$beta_hat)), 
           method.name = c('MuSiC', 'NNLS', 'equal-weight-smooth','equal-weight-local','music-weight-smooth','music-weight-local'))
```

## Enge as bulk, Seger as reference

create pseudo-bulk data using Enge dataset such that minimum prop is 0.2


```{r}
enge_bulk = create_bulk(enge,indis = levels(factor(enge$individual))[-8],cell_types =  c("acinar",'alpha','beta','ductal'),min_prop = 0.2)
enge_bulk
```


try music using enge dataset

```{r}
bulk.data = ExpressionSet(assayData = counts(enge_bulk$bulk))
#pheno_Data = data.frame(cellType = seger$cell_type,SubjectName = seger$individual)
#rownames(pheno_Data) = colnames(seger)
#ref.data = ExpressionSet(assayData = counts(seger),phenoData = AnnotatedDataFrame(pheno_Data))

Est.prop.enge = music_prop(bulk.eset = bulk.data, sc.eset = seger.ref,
                           clusters = 'cellType', samples = 'SubjectName', select.ct = c("acinar",'alpha','beta','ductal'))
Eval_multi(prop.real = data.matrix(enge_bulk$bulk_prop), 
           prop.est = list(data.matrix(Est.prop.enge$Est.prop.weighted), 
                           data.matrix(Est.prop.enge$Est.prop.allgene)), 
           method.name = c('MuSiC', 'NNLS'))

round(enge_bulk$bulk_prop,2)
round(Est.prop.enge$Est.prop.weighted,2)
```

Not very precise estimation when using all common genes. Now try to random select some genes to create a "good case".



```{r}
set.seed(01)
cm.gene = intersect(rownames(seger.ref),rownames(bulk.data))
cm.gene = cm.gene[sample(1:length(cm.gene),3000)]
g.bulk = match(cm.gene,rownames(bulk.data))
g.ref = match(cm.gene,rownames(seger.ref))

Est.prop.enge = music_prop(bulk.eset = bulk.data[g.bulk,], sc.eset = seger.ref[g.ref,],
                           clusters = 'cellType', samples = 'SubjectName', select.ct = c("acinar",'alpha','beta','ductal'))
Eval_multi(prop.real = data.matrix(enge_bulk$bulk_prop), 
           prop.est = list(data.matrix(Est.prop.enge$Est.prop.weighted), 
                           data.matrix(Est.prop.enge$Est.prop.allgene)), 
           method.name = c('MuSiC', 'NNLS'))
round(Est.prop.enge$Est.prop.weighted,2)
round(Est.prop.enge$Est.prop.allgene,2)
```




run deconference

```{r, results='asis'}
datax = set_data_decon(exprs(bulk.data),seger,marker_gene = rownames(Est.prop.enge$Weight.gene),
                       cell_types = c("acinar",'alpha','beta','ductal'), gene_thresh = 0,max_count_quantile = 1,w=1)
out5 = deconference(datax,est_pop_var = TRUE,meta_mode = 'smooth',correction = T,verbose = TRUE,hc.type = 'hc3')
summary_real(out5,as.matrix(enge_bulk$bulk_prop))

out6 = deconference(datax,est_pop_var = TRUE,meta_mode = 'local',correction = T,verbose = TRUE,hc.type = 'hc3')
summary_real(out6,as.matrix(enge_bulk$bulk_prop))
```

```{r}
round(Est.prop.enge$S,2)
round(out5$input$S,2)
```

```{r, results='asis'}
w = Est.prop.enge$Weight.gene
w = apply(w,2,function(z){z/sum(z,na.rm = T)})
w = rowMeans(w,na.rm = T)
datax = set_data_decon(exprs(bulk.data),seger,marker_gene = rownames(Est.prop.enge$Weight.gene),
                       cell_types = c("acinar",'alpha','beta','ductal'), gene_thresh = 0,max_count_quantile = 1,w=w)
out7 = deconference(datax,est_pop_var = TRUE,meta_mode = 'smooth',correction = T,verbose = TRUE,hc.type = 'hc3')
summary_real(out7,as.matrix(enge_bulk$bulk_prop))

out8 = deconference(datax,est_pop_var = TRUE,meta_mode = 'local',correction = T,verbose = TRUE,hc.type = 'hc3')
summary_real(out8,as.matrix(enge_bulk$bulk_prop))
```


```{r, results='asis'}
datax = set_data_decon(exprs(bulk.data),seger,marker_gene = rownames(Est.prop.enge$Weight.gene),
                       cell_types = c("acinar",'alpha','beta','ductal'), gene_thresh = 0,max_count_quantile = 1,w=NULL)
out9 = deconference(datax,est_pop_var = TRUE,meta_mode = 'smooth',correction = T,verbose = TRUE,hc.type = 'hc3')
summary_real(out9,as.matrix(enge_bulk$bulk_prop))

out10 = deconference(datax,est_pop_var = TRUE,meta_mode = 'local',correction = T,verbose = TRUE,hc.type = 'hc3')
summary_real(out10,as.matrix(enge_bulk$bulk_prop))
```


```{r}
Eval_multi(prop.real = data.matrix(enge_bulk$bulk_prop), 
           prop.est = list(data.matrix(Est.prop.enge$Est.prop.weighted), 
                           data.matrix(Est.prop.enge$Est.prop.allgene),
                           t(out5$beta_hat),
                           t(out6$beta_hat),
                           t(out7$beta_hat),
                           t(out8$beta_hat),
                           t(out9$beta_hat),
                           t(out10$beta_hat)), 
           method.name = c('MuSiC', 'NNLS', 'equal-weight-smooth','equal-weight-local','music-weight-smooth',
                           'music-weight-local','default-weight-smooth','default-weight-local'))
```
