---
title: "oracle mse xin data"
author: "DongyueXie"
date: "2020-12-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

We have matrices $X_i$ for each individual $i = 1,2,...,n$. 

Let $y_i = X_i\beta_i$ and $X = \frac{1}{n}\sum_i X_i$, we fit the following model to estimate $\beta_i$:

\[y_i  =  X\beta_i +(X_i-X)\beta_i:=X\beta_i + \epsilon_i\]

We can estimate the variance of $\epsilon_i$ as the sample covariance of $(X_i-X)\beta_i$ and we denote the estimated one as $\Sigma_i$.

Thus, the estimator is $\hat\beta_i = (X^T\Sigma_i^{-1}X)^{-1}X^T\Sigma_i^{-1}y$



```{r}
X_xin <- readRDS("data/pancreas/X_xin.rds")
X_array = X_xin$X_array
X_array = X_array[, , -c(4, 7)]

N = dim(X_array)[1]
K = dim(X_array)[2]
NI = dim(X_array)[3]

rm.gene = which(rowSums(rowSums(X_array,dims=2))==0)

for(k in 1:K){
  gene_mat_k = X_array[,k,]
  rm.gene = c(rm.gene,which(rowSums(gene_mat_k)==0))
}


X_array = X_array[-rm.gene,,]

X = rowMeans(X_array,dims = 2)

b = c(0.2,0.3,0.5)
y = apply(X_array,3,function(z){z%*%b})

Sigma_sample = c()
for(i in 1:16){
  Sigma_sample = cbind(Sigma_sample,(X_array[,,i] - X)%*%b)
}

Sigma = cov(t(Sigma_sample))
Sigma_inv = MASS::ginv(Sigma)

simu_result_oracle = function(Xs,beta,Sigma_inv){
  
  y <- apply(Xs, 3, function(mat) mat %*% beta)
  X <- rowMeans(Xs, dims = 2)

  V = solve(t(X)%*%Sigma_inv%*%X)
  A = V%*%t(X)%*%Sigma_inv

  std.sd <- sqrt(diag(V))

est.beta <- apply(y, 2, function(v) {
  b_hat = A%*%v
  temp <- pmax(b_hat, 0)
  return((c(temp/sum(temp), (temp - beta) / std.sd, std.sd)))
})
rownames(est.beta) <- c("Est1", "Est2", "Est3", "dev1.LS", "dev2.LS", "dev3.LS", "sd1.LS", "sd2.LS", "sd3.LS")
MSE <- sqrt(colSums((est.beta[1:3, ] - beta)^2))
#print(rbind(est.beta, MSE))
rbind(est.beta, MSE)
}

out_oracle = simu_result_oracle(X_array,b,Sigma_inv)
round(out_oracle,3)
```

It does not work. Maybe the number of individual is too small. 


How about transform $\Sigma$ to correlation matrix.

```{r}
library(sandwich)
# obj, a lm fitted model
# Omega, correlation materix
vcovHCC = function(obj,Omega=NULL,hc.type='hc3'){
  library(Rfast)
  X = as.matrix(obj$model[,-1])
  if(is.null(Omega)){
    A = mat.mult(mat.mult(solve(Crossprod(X,X)),t(X)),as.matrix(obj$residuals))
    V = Tcrossprod(A,A)
  }else{
    A = mat.mult(solve(Crossprod(X,X)),t(X))
    d = abs(obj$residuals)
    if(hc.type=='hc3'){
      d = d/(1-influence(obj)$hat)
    }
    Sigma = t(t(Omega*d)*d)
    V = mat.mult(mat.mult(A,Sigma),t(A))
  }
  V
}



simu_result = function(X,beta,Omega){
  
  y <- apply(X, 3, function(mat) mat %*% beta)
mu <- rowMeans(X, dims = 2)

rm.idx = which(rowSums(mu)==0)


est.beta <- apply(y, 2, function(v) {
  temp <- lm(v ~ mu + 0)
  sw.sd <- sqrt(diag(vcovHC(temp, type="HC3")))
  std.sd <- sqrt(diag(vcov(temp)))
  swc.sd <- sqrt(diag(vcovHCC(temp,Omega=Omega)))
  temp <- pmax(temp$coef, 0)
  return((c(temp/sum(temp), (temp - beta) / std.sd, (temp - beta) / sw.sd, (temp - beta) / swc.sd,std.sd,sw.sd,swc.sd)))
})
rownames(est.beta) <- c("Est1", "Est2", "Est3", "dev1.OLS", "dev2.OLS", "dev3.OLS", "dev1.sandwich", "dev2.sandwich", "dev3.sandwich","dev1.sandwich.cor", "dev2.sandwich.cor", "dev3.sandwich.cor","sd1.OLS", "sd2.OLS", "sd3.OLS", "sd1.sandwich", "sd2.sandwich", "sd3.sandwich","sd1.sandwich.cor", "sd2.sandwich.cor", "sd3.sandwich.cor")
MSE <- sqrt(colSums((est.beta[1:3, ] - beta)^2))
#print(rbind(est.beta, MSE))
rbind(est.beta, MSE)
}

out = simu_result(X_array,b,cov2cor(Sigma))

round(out,3)
```

```{r}
hist(cov2cor(Sigma),breaks = 1000)
```

As a comparison, plot the gene correlation estiamted from raw count data.

```{r}
gene_cor = readRDS("data/pancreas/gene_cor.rds")
gene_cor = gene_cor[-rm.gene,-rm.gene]
hist(gene_cor,breaks = 1000)
```



