---
title: "real data benchmark"
author: "DongyueXie"
date: "2020-08-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F,message = F)
```

## Introduction

Human pancreas real data: Xin et al 2016, Segerstolpe et al. (2016), Muraro et al (2016), Baron et al(2016). 

Xin et al: 1492 cells, 4 cell types, 18(12 healthy + 6 T2D) individuals, Illumina HiSeq 2500.

Segerstolpe et al: 2209 cells, 14+1 cell types, 10(6 healthy + 4 T2D) individuals, Smart-seq2.

Muraro et al: 2126 cells, 14 cell types, 4 donors, CEL-seq2.

Baron et al: 8569 cells, 14 cell types, 4 cadaveric donors, inDrop.

```{r,echo=F}
summary_real = function(out,true_b){
  
  # compare fitted vs true
  b_hat = t(out$beta_hat)
  rownames(b_hat) = rownames(true_b)
  colnames(b_hat) = colnames(true_b)
  print(knitr::kable(
  list(t0 = round(b_hat,2), t1 = round(true_b,2)),
  caption = 'estimated vs true beta',
  booktabs = TRUE, valign = 't'))
  
  
  ci_length = t(round(out$beta_se,2))
  colnames(ci_length) = colnames(true_b)
  rownames(ci_length) = rownames(true_b)
  print(knitr::kable(ci_length,caption = 'beta hat se'))
  
  # summary statistics
  rmse = sqrt(mean((t(out$beta_hat)-true_b)^2))
  mad = mean(abs(t(out$beta_hat)-true_b))
  r2 = cor(c(t(out$beta_hat)),c(true_b))
  ss = rbind(c(rmse,mad,r2))
  colnames(ss) = c('rmse','mad','r2')
  rownames(ss) = NULL
  
  print(knitr::kable(ss,caption='error measure'))
  
  # coverage, length of CI
  ci = build_ci(out)
  coverage = c(true_b>=t(ci$ci_l))&(true_b<=t(ci$ci_r))
  
  print(knitr::kable(coverage,caption='coverage'))

  # two group test
  
  groups = c(1,3,1,1,2,3,2,2,3,3)
  g1 = which(groups==1)
  g2 = which(groups==2)
  test2 = two_group_test(out,groups)
  
  test_out = rbind(round(colMeans(true_b[g1,]) - colMeans(true_b[g2,]),3),
                   round(test2$diff_group,3),
                   round(sqrt(diag(test2$V_tilde)),3),
                   round(test2$p_value,3))
  rownames(test_out) = c('true diff','estimated diff','se','p value')
  print(knitr::kable(test_out,caption= ' two group test'))
  
}
```


## Xin et al 2016

Total 18 individuals. I took the first 8 healthy individuals as reference data and aggregate cells for the rest 10 individuals(4 healthy and 6 T2D) as reference data. 

two group test: Non T2D 9,11,12 vs T2D 1,3,4

weight: default; 

```{r, results='asis'}
library(MuSiC)
library(xbioc)
source('code/deconference_main.R')
XinT2Deset <- readRDS("data/MuSiC/XinT2Deset.rds")

XinT2D.construct.full = bulk_construct(XinT2Deset, clusters = 'cellType', samples = 'SubjectName')
XinT2D.construct.full$prop.real = relative.ab(XinT2D.construct.full$num.real, by.col = FALSE)
knitr::kable(list(t0=table(XinT2Deset$SubjectName, XinT2Deset$cellType),t1=round(as.matrix(XinT2D.construct.full$prop.real),2)),
             caption = 'cell counts and proportinos in each individual',
             booktabs = TRUE, valign = 't')
dim(exprs(XinT2Deset))
hist(colMeans(exprs(XinT2Deset)),breaks=100,main='cell library size / #gene',xlab = 'size')
true_b = as.matrix(XinT2D.construct.full$prop.real[-c(1:8),])
```

### Universal 

**HC3**

```{r}
load('output/Xin_universal_hc3.RData')
dim(out_universal$input$X)
out_universal$input$Sigma[1,1] * nrow(out_universal$input$X)^2
```

```{r, results='asis'}
summary_real(out_universal,true_b)
```

**HC2**

```{r, results='asis'}
load('output/Xin_universal_hc2.RData')
summary_real(out_universal_hc2,true_b)
```

**HC0**

```{r, results='asis'}
load('output/Xin_universal_hc0.RData')
summary_real(out_universal_hc0,true_b)
```


### none-meta analysis

**HC3**

```{r, results='asis'}
load('output/Xin_sampleM_hc3.RData')
summary_real(out_sampleM_hc3,true_b)
```


### unadjusted

```{r, results='asis'}
summary_real2 = function(b_hat,beta_se,true_b){
  
  # compare fitted vs true
  b_hat = t(b_hat)
  rownames(b_hat) = rownames(true_b)
  colnames(b_hat) = colnames(true_b)
  print(knitr::kable(
  list(t0 = round(b_hat,2), t1 = round(true_b,2)),
  caption = 'estimated vs true beta',
  booktabs = TRUE, valign = 't'))
  
  
  ci_length = t(round(beta_se,2))
  colnames(ci_length) = colnames(true_b)
  rownames(ci_length) = rownames(true_b)
  print(knitr::kable(ci_length,caption = 'beta hat se'))
  
  # summary statistics
  rmse = sqrt(mean((b_hat-true_b)^2))
  mad = mean(abs(b_hat-true_b))
  r2 = cor(c(b_hat),c(true_b))
  ss = rbind(c(rmse,mad,r2))
  colnames(ss) = c('rmse','mad','r2')
  rownames(ss) = NULL
  
  print(knitr::kable(ss,caption='error measure'))
  
  # coverage, length of CI
  coverage = c(true_b>=(b_hat-2*t(beta_se)))&(true_b<=(b_hat+2*t(beta_se)))
  
  print(knitr::kable(coverage,caption='coverage'))

  
}

load('output/Xin_sampleM_unadj.RData')
summary_real2(out_unadj$beta_hat,out_unadj$sand.out$beta_se,true_b)
summary_real2(out_unadj$beta_hat,out_unadj$sand.out.hc3$beta_se,true_b)
```




```{r,eval=F}
## Segerstolpe et al

segerstolpe <- readRDS("data/MuSiC/segerstolpe.rds")
table(segerstolpe$age[which(segerstolpe$cell_type1%in%c('alpha','beta','delta','gamma'))],
      segerstolpe$cell_type1[which(segerstolpe$cell_type1%in%c('alpha','beta','delta','gamma'))])
table(segerstolpe$disease[which(segerstolpe$cell_type1%in%c('alpha','beta','delta','gamma'))],
      segerstolpe$cell_type1[which(segerstolpe$cell_type1%in%c('alpha','beta','delta','gamma'))])
table(segerstolpe$age,
      segerstolpe$disease)

dim(segerstolpe@assays@.xData$data$counts)
hist(colMeans((segerstolpe@assays@.xData$data$counts)),breaks = 100,main='cell library size / #gene',xlab = 'size')
```
