---
title: "real data simulation"
author: "DongyueXie"
date: "2020-10-29"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Random select a half individuals from each dataset and create bulk datasets, half of wiich are assigned to healthy groups and the other half to disease group.

For reference dataset, 1. random select a half of rest half data from each dataset as reference data; 2. use all of the rest dataset.

issues: add cell size estimate? add gene length adjustment?



```{r,message=FALSE,warning=FALSE}
library(MuSiC)
library(Biobase)
library(xbioc)
source('code/deconference_main.R')
# #load datasets
# # 4
# baron <- readRDS("data/pancreas/baron.rds")
# # 4
# muraro <- readRDS("data/pancreas/muraro.rds")
# # 10
# seger <- readRDS("data/pancreas/segerstolpe_raw.rds")
# # 8
# enge = readRDS("data/pancreas/enge_raw.rds")
# # 18
# xin = readRDS('data/pancreas/xin_raw.rds')
refs_raw = readRDS('data/pancreas/refs_raw.rds')
```

```{r}

summary_real = function(out,true_b,groups){
  
  # compare fitted vs true
  b_hat = t(out$beta_hat)
  rownames(b_hat) = rownames(true_b)
  colnames(b_hat) = colnames(true_b)
  print(knitr::kable(
  list(t0 = round(b_hat,2), t1 = round(true_b,2)),
  caption = 'estimated vs true beta',
  booktabs = TRUE, valign = 't'))
  
  
  ci_length = t(round(out$beta_se,2))
  colnames(ci_length) = colnames(true_b)
  rownames(ci_length) = rownames(true_b)
  print(knitr::kable(ci_length,caption = 'beta hat se'))
  
  # summary statistics
  rmse = sqrt(mean((t(out$beta_hat)-true_b)^2))
  mad = mean(abs(t(out$beta_hat)-true_b))
  r2 = cor(c(t(out$beta_hat)),c(true_b))
  ss = rbind(c(rmse,mad,r2))
  colnames(ss) = c('rmse','mad','r2')
  rownames(ss) = NULL
  
  print(knitr::kable(ss,caption='error measure'))
  
  # coverage, length of CI
  ci = build_ci(out)
  coverage = c(true_b>=t(ci$ci_l))&(true_b<=t(ci$ci_r))
  
  print(knitr::kable(coverage,caption='coverage'))

  # two group test
  group_name = levels(as.factor(groups))
  g1 = which(groups==group_name[1])
  g2 = which(groups==group_name[2])
  test2 = two_group_test(out,groups)

  test_out = rbind(round(colMeans(true_b[g1,]) - colMeans(true_b[g2,]),3),
                   round(test2$diff_group,3),
                   round(sqrt(diag(test2$V_tilde)),3),
                   round(test2$p_value,3))
  rownames(test_out) = c('true diff','estimated diff','se','p value')
  print(knitr::kable(test_out,caption= ' two group test'))
  
}


# data.obj is a list of single cell dataset
create_realsimu_data = function(refs,cell_types,proph,propd,ncells = 1000,indis_ref = NULL){
  G = nrow(refs[[1]])
  bulk_counts = matrix(0,nrow=G,ncol=50)
  bulk_prop = c()
  all_indis = c()
  condition = c()
  indi_counter = 0
  ref1 = list()
  ref2 = list()
  for(d in 1:length(refs)){
    
    # random sample two individuals
    data.obj = refs[[d]]
    
    ## these indis are for bulk data
    if(is.null(indis_ref)){
      c.idx = which(data.obj$cell_type%in%cell_types)
      temp = table(data.obj$individual[c.idx],data.obj$cell_type[c.idx])
      temp = temp[,match(cell_types,colnames(temp))]
      valid_indi = which(rowSums(temp==0)==0)
      indis = sample(rownames(temp)[valid_indi],ceiling(length(valid_indi)/2/2)*2)
    }else{
      indis = indis_ref[[d]]
    }
    all_indis = c(all_indis,indis)
    condition = c(condition,c(rep('health',length(indis)/2),rep('disease',length(indis)/2)))
    #print(indis)
    # first half health
    # second half disease
    for(i in 1:(length(indis)/2)){
      idx.temp = which(data.obj$individual==indis[i]&data.obj$cell_type%in%cell_types)
      cell_counts = table(data.obj$cell_type[idx.temp])
      if(is.null(ncells)){
      ncells = round(max(cell_counts/proph))
      }
      bcounts = 0
      for(j in 1:length(cell_types)){
        idx_ij = which(data.obj$individual==indis[i]&data.obj$cell_type==cell_types[j])
        #print(idx_ij)
        #print(data.obj[,sample(idx_ij,ncells*prop[j],replace = TRUE)])
        bcounts = bcounts + rowSums(counts(data.obj[,sample(idx_ij,ncells*proph[j],replace = TRUE)]))
      }
      bulk_prop = rbind(bulk_prop,proph)
      indi_counter = indi_counter + 1
      bulk_counts[,indi_counter] = bcounts
      
    }
    
  
  ## disease
  for(i in (length(indis)/2+1):(length(indis))){
      idx.temp = which(data.obj$individual==indis[i]&data.obj$cell_type%in%cell_types)
      cell_counts = table(data.obj$cell_type[idx.temp])
      if(is.null(ncells)){
      ncells = round(max(cell_counts/propd))
      }
      bcounts = 0
      for(j in 1:length(cell_types)){
        idx_ij = which(data.obj$individual==indis[i]&data.obj$cell_type==cell_types[j])
        #print(idx_ij)
        #print(data.obj[,sample(idx_ij,ncells*prop[j],replace = TRUE)])
        bcounts = bcounts + rowSums(counts(data.obj[,sample(idx_ij,ncells*propd[j],replace = TRUE)]))
      }
      bulk_prop = rbind(bulk_prop,propd)
      indi_counter = indi_counter + 1
      bulk_counts[,indi_counter] = bcounts
  }
    
    
    ## the rest indis are for reference data
    
    ### the first reference data uses half the rest data
    rest_indis = unique(data.obj$individual)[-match(indis,unique(data.obj$individual))]
    ref1_indis = sample(rest_indis,length(rest_indis)/2)
    ref1_idx = which(data.obj$individual%in%ref1_indis&data.obj$cell_type%in%cell_types)
    ref1[[d]] = data.obj[,ref1_idx]
    
    ### the second reference data uses all of the half dataset
    ref2_idx = which(data.obj$individual%in%rest_indis&data.obj$cell_type%in%cell_types)
    ref2[[d]] = data.obj[,ref2_idx]
}

  #browser()
  bulk_counts = bulk_counts[,which(colSums(bulk_counts)!=0)]
  colnames(bulk_counts) = all_indis
  rownames(bulk_counts) = rownames(data.obj)
  #rownames(bulk_ncell) = indis
  rownames(bulk_prop) = all_indis
  colnames(bulk_prop) = cell_types
  #ii = which(data.obj$individual%in%indis&data.obj$cell_type%in%cell_types)
  #bulk_prop = table(data.obj$individual[ii],data.obj$cell_type[ii])
  #bulk_prop = bulk_prop/rowSums(bulk_prop)
  coldata = DataFrame(individual = all_indis,condition=condition)
  #browser()
  out = SingleCellExperiment(assays = list(counts = as.matrix(bulk_counts)),
                                   colData = coldata)
  return(list(bulk = out,bulk_prop = bulk_prop,group = condition,ref1=ref1,ref2=ref2))
}

sce_to_es = function(sce){
  #bulk.data = ExpressionSet(assayData = counts(enge_bulk$bulk))
  pheno_Data = data.frame(cell_type = sce$cell_type,individual = sce$individual)
  rownames(pheno_Data) = colnames(sce)
  ref.data = ExpressionSet(assayData = counts(sce),phenoData = AnnotatedDataFrame(pheno_Data))
  ref.data
}
```


```{r}
set.seed(12345)
cell_types = c('alpha','beta','delta')
datax = create_realsimu_data(refs_raw,cell_types,c(0.3,0.5,0.2),c(0.3,0.2,0.5))
```

now we have create bulk data belonging to two groups - health and disease, and the reference data set.

## ref1

```{r}

out_ref1 = deconference_multi_ref(datax$ref1,datax$bulk,cell_types = cell_types,verbose = F,cellsize_est = 'ols')
out_ref1_nondiag = deconference_multi_ref(datax$ref1,datax$bulk,cell_types = cell_types,verbose = F,cellsize_est = 'ols',est_sigma2 = F)

merge_ref = function(refs){
  
  ref1_es = refs[[1]]
colData(ref1_es) = DataFrame(cell_type = ref1_es$cell_type,individual = ref1_es$individual)
rowData(ref1_es) = DataFrame(feature_symbol = rowData(ref1_es)$feature_symbol)
#colnames(ref1_es) = ref1_es$cell_type
reducedDims(ref1_es) = NULL
assays(ref1_es)$logcounts = NULL
for(i in 2:length(datax$ref1)){
  temp = refs[[i]]
  colData(temp) = DataFrame(cell_type = temp$cell_type,individual = temp$individual)
  rowData(temp) = DataFrame(feature_symbol = rowData(temp)$feature_symbol)
  #colnames(temp) = temp$cell_type
  reducedDims(temp) = NULL
  if(!is.null(assays(temp)$logcounts)){
    assays(temp)$logcounts = NULL
  }
  ref1_es = cbind(ref1_es,temp)
}
ref1_es

}

ref1_es = merge_ref(datax$ref1)

bulk_es = ExpressionSet(assayData = counts(datax$bulk))
ref1_es = sce_to_es(ref1_es)
music_out1 = music_prop(bulk_es,ref1_es,clusters = 'cell_type',samples = 'individual',verbose = F)
Eval_multi(prop.real = data.matrix(datax$bulk_prop), 
           prop.est = list(data.matrix(music_out1$Est.prop.weighted),
                           t(out_ref1$beta_hat),
                           t(out_ref1_nondiag$beta_hat)), 
           method.name = c('MuSiC','deconinfer','deconinfer-nondiag'))

```

```{r,results='asis'}
summary_real(out_ref1,datax$bulk_prop,datax$group)
summary_real(out_ref1_nondiag,datax$bulk_prop,datax$group)
```

## ref2 

```{r}
out_ref2 = deconference_multi_ref(datax$ref2,datax$bulk,cell_types = cell_types,verbose = F,cellsize_est = 'ols')
out_ref2_nondiag = deconference_multi_ref(datax$ref2,datax$bulk,cell_types = cell_types,verbose = F,cellsize_est = 'ols',est_sigma2 = F)
ref2_es = merge_ref(datax$ref2)

ref2_es = sce_to_es(ref2_es)
music_out2 = music_prop(bulk_es,ref2_es,clusters = 'cell_type',samples = 'individual')
Eval_multi(prop.real = data.matrix(datax$bulk_prop), 
           prop.est = list(data.matrix(music_out2$Est.prop.weighted),t(out_ref2$beta_hat),
                           t(out_ref1_nondiag$beta_hat)), 
           method.name = c('MuSiC','deconinfer','deconinfer-nondiag'))
```

```{r,results='asis'}
summary_real(out_ref2,datax$bulk_prop,datax$group)
summary_real(out_ref2_nondiag,datax$bulk_prop,datax$group)
```



# adjust gene length

```{r}
refs = readRDS('data/pancreas/refs.rds')
```

```{r}
set.seed(12345)
cell_types = c('alpha','beta','delta')
datax = create_realsimu_data(refs,cell_types,c(0.2,0.6,0.2),c(0.2,0.2,0.6))
```

now we have create bulk data belonging to two groups - health and disease, and the reference data set.

## ref1

```{r}
out_ref1 = deconference_multi_ref(datax$ref1,datax$bulk,cell_types = cell_types,verbose = F,cellsize_est = 'ols')
out_ref1_nondiag = deconference_multi_ref(datax$ref1,datax$bulk,cell_types = cell_types,verbose = F,cellsize_est = 'ols',est_sigma2 = F)



ref1_es = merge_ref(datax$ref1)

bulk_es = ExpressionSet(assayData = counts(datax$bulk))
ref1_es = sce_to_es(ref1_es)
music_out1 = music_prop(bulk_es,ref1_es,clusters = 'cell_type',samples = 'individual')
Eval_multi(prop.real = data.matrix(datax$bulk_prop), 
           prop.est = list(data.matrix(music_out1$Est.prop.weighted),
                           t(out_ref1$beta_hat),
                           t(out_ref1_nondiag$beta_hat)), 
           method.name = c('MuSiC','deconinfer','deconinfer-nondiag'))
```

```{r,results='asis'}
summary_real(out_ref1,datax$bulk_prop,datax$group)
summary_real(out_ref1_nondiag,datax$bulk_prop,datax$group)
```

## ref2 

```{r}
out_ref2 = deconference_multi_ref(datax$ref2,datax$bulk,cell_types = cell_types,verbose = F,cellsize_est = 'ols')
out_ref2_nondiag = deconference_multi_ref(datax$ref2,datax$bulk,cell_types = cell_types,verbose = F,cellsize_est = 'ols',est_sigma2 = F)
ref2_es = merge_ref(datax$ref2)

ref2_es = sce_to_es(ref2_es)
music_out2 = music_prop(bulk_es,ref2_es,clusters = 'cell_type',samples = 'individual')
Eval_multi(prop.real = data.matrix(datax$bulk_prop), 
           prop.est = list(data.matrix(music_out2$Est.prop.weighted),t(out_ref2$beta_hat),
                           t(out_ref1_nondiag$beta_hat)), 
           method.name = c('MuSiC','deconinfer','deconinfer-nondiag'))

```

```{r,results='asis'}
summary_real(out_ref2,datax$bulk_prop,datax$group)
summary_real(out_ref2_nondiag,datax$bulk_prop,datax$group)
```

