---
title: "sandwich estimator error in variable and generalized least square"
author: ""
date: "2020-07-02"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F)
```


In a previous [simulation study](sandwich_error.html), we showed that HC3 could give estimates closer to the true variances, and the Jackknife gives very similar estimates to HC3, though their equivalence is not straightforward. Here, we design a simulation study to examine: 1. Does HC3 and Jackknife still be similar if we increase error in variable; 2. How about 1 and add weights? If adding weights, then the Jackknife estimate of variance is not consistent?; 3. How about gls estimator? Add estimated weights; 4. Add random effect term see if it could correct for bias. 

We generate data from \[y_i\sim Poisson(x_i^T\beta), x_i\sim Uniform(a,b), X_i = x_i+v_i, v_i\sim N(0,\Sigma)\] 

In the model, $x_i$ is the true regressor and $X_i$ is the one with error in it. We can tune $a,b,\Sigma$ such that the variations in $X$ are different and see how methods perform.

```{r}
source('code/sandwich_var.R')
simu_study_sandwich = function(n,p,b,a,Sigma=diag(p),x=NULL,add_weights=FALSE,optimal=FALSE,correction = TRUE,nreps=100){
  
  beta_hats = matrix(nrow=nreps,ncol=p)
  beta_ses = matrix(nrow=nreps,ncol=p*4)
  
  if(is.null(x)){
    x = t(gtools::rdirichlet(p,rep(a,n)))*n*10
  }else{
    x = apply(x,2,function(z){z/sum(z)})
    x = x * n * 10
  }
  
  
  for(i in 1:nreps){
  
  
  y = rpois(n,x%*%b)
  X = x + MASS::mvrnorm(n,rep(0,p),Sigma)
  if(add_weights){
    if(optimal){
      w = c(1/(x%*%b))
    }else{
      w = c(1/(rowSums(X)))
    }
    
  }else{
    w = NULL
  }
  
  
  
  
  #hc0
  beta_se_hc0 = sqrt(diag(sandwich_var(y,X,Sigma,w,type='hc',correction=correction)$asyV))
  #hc2
  beta_se_hc2 = sqrt(diag(sandwich_var(y,X,Sigma,w,type='hc2',correction=correction)$asyV))
  #hc3
  beta_se_hc3 = sqrt(diag(sandwich_var(y,X,Sigma,w,type='hc3',correction=correction)$asyV))
  #jack
  beta_se_jack = sqrt(diag(jackknife_var(y,X,Sigma,w)))
  
  beta_hats[i,] = sandwich_var(y,X,Sigma,w,type='hc',correction=correction)$b_hat
  beta_ses[i,] = c(beta_se_hc0,beta_se_hc2,beta_se_hc3,beta_se_jack)
  }

   return(list(b=b,beta_hats=beta_hats,beta_ses=beta_ses))
  
}

simu_study_summary = function(out){
  
  beta_hats = out$beta_hats
  beta_ses = out$beta_ses
  b = out$b
  nreps = nrow(out$beta_hats)
  
  beta_hat_se <- matrix(colMeans(beta_ses), nrow = 4, byrow = T)
  beta_hat_se <- rbind(beta_hat_se, apply(beta_hats,2,sd))
  rownames(beta_hat_se) <- c("Sandwich","HC2","HC3","Jackknife","True")
  colnames(beta_hat_se) = 1:length(b)
  print(knitr::kable(round(beta_hat_se,4),caption="compare mean of estiamted variance"))



  true_betas = do.call('cbind',rep(list(rep(1,nreps)%*%t(b)),4))
  beta_hats_rep = do.call('cbind',rep(list(beta_hats),4))

  coverage = (true_betas>=beta_hats_rep-1.96*beta_ses)&(true_betas<=beta_hats_rep+1.96*beta_ses)
  coverage <- matrix(colMeans(coverage), nrow = 4, byrow = T)
  rownames(coverage) <- c("Sandwich","HC2","HC3","Jackknife")
  colnames(coverage) = 1:length(b)
  print(knitr::kable(round(coverage,3),caption="compare coverages of beta"))

}

```

Let's start with no error in variable.

```{r}
set.seed(12345)
out = simu_study_sandwich(n=100,p=3,b=1:3/6,a=0.5,Sigma = 0*diag(3),nreps = 300,correction = F)
simu_study_summary(out)
```

Add noise to x

```{r}
set.seed(12345)
out = simu_study_sandwich(n=100,p=3,b=1:3/6,a=0.5,Sigma = 100*diag(3),nreps = 300,correction = F)
simu_study_summary(out)
```


```{r}
set.seed(12345)
out = simu_study_sandwich(100,3,1:3/6,a=0.5,Sigma = 200*diag(3),nreps = 300,correction = F)
simu_study_summary(out)
```


```{r}
set.seed(12345)
out = simu_study_sandwich(100,3,1:3/6,a=0.5,Sigma = 300*diag(3),nreps = 300,correction = F)
simu_study_summary(out)
```


Add weights


```{r}
set.seed(12345)
out = simu_study_sandwich(n=100,p=3,b=1:3/6,a=0.5,Sigma = 0*diag(3),nreps = 300,add_weights = TRUE,optimal = TRUE,correction = FALSE)
simu_study_summary(out)
```


```{r}
set.seed(12345)
out = simu_study_sandwich(n=100,p=3,b=1:3/6,a=0.5,Sigma = 100*diag(3),nreps = 300,add_weights = TRUE,optimal = TRUE,correction = FALSE)
simu_study_summary(out)
```


```{r}
set.seed(12345)
out = simu_study_sandwich(n=100,p=3,b=1:3/6,a=0.5,Sigma = 100*diag(3),nreps = 300,add_weights = TRUE,optimal = TRUE,correction = TRUE)
simu_study_summary(out)
```


