---
title: "Check variance estimate and normality in simulation, when correlation presents, simulated U"
author: "Dongyue Xie"
date: "2021-06-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Simulated 1000 gene, 4 cell type, 4 bulk individuals

```{r,warning=F,message=F}
source('code/deconference_main.R')
source('code/utils.R')
source('code/wols.R')
source('code/simulation/simu_correlation.R')

ploter.temp = function(results){

  par(mfrow=c(2,2))

  plot(results$coverage_adj_hc0,
       ylim = range(c(results$coverage_adj_hc0,results$coverage_unadj_hc0,results$coverage_unadj_cv,1),na.rm = T),
       col=2,ylab='coverage',xlab='coefs',main='coverage of p',type='b')
  lines(results$coverage_unadj_hc0,type='b',pch=2,col=4)
  lines(results$coverage_unadj_cv,type='b',pch=2,col=3)
  abline(h=0.95,lty=2)

  plot(results$coverage_diff_adj_hc0,ylim = range(c(results$coverage_diff_adj_hc0,results$coverage_diff_unadj_hc0,results$coverage_diff_unadj_cv,1),na.rm = T),
       col=2,type='b',ylab='coverage',xlab='coefs',main='coverage of difference')
  lines(results$coverage_diff_unadj_hc0,type='b',pch=2,col=4)
  lines(results$coverage_diff_unadj_cv,type='b',pch=2,col=3)
  abline(h=0.95,lty=2)

  #p_order = order(abs(results$p-results$p2))

  plot(results$power_adj_hc0,ylim = range(c(results$power_adj_hc0,results$power_unadj_hc0,results$power_unadj_cv),na.rm = T),
       col=2,ylab="power",xlab='',main='power')
  lines(results$power_unadj_hc0,type='p',pch=2,col=4)
  lines(results$power_unadj_cv,type='p',pch=2,col=3)
  abline(h=0.05,lty=2)

  legend('bottomright',c('adjusted_hc0','unadj_hc0',"unadj_const"),col=c(2,4,3),pch=c(1,2,2))

  par(mfrow=c(1,1))

}
```

```{r}
G = 1000
K = 4
set.seed(12345)
ref = matrix(rnorm(G*K),nrow=G)
ref = abs(ref)
ref = apply(ref, 2, function(z){z/sum(z)})*G
sigma2 = ref/2
rownames(ref) = 1:G
b1 = c(0.1,0.1,0.3,0.5)
b2 = c(0.1,0.2,0.5,0.2)
nb = 4
b.m = cbind(b1%*%t(rep(1,nb/2)),b2%*%t(rep(1,nb/2)))
```

```{r}
p = G
A = matrix(0,nrow=p,ncol=p)

for(i in 1:p){
  for(j in 1:p){
    A[i,j] = max(1-abs(i-j)/50,0)
  }
}
```


```{r}
set.seed(12345)
simu = simu_corr_simple(ref,b.m,nreps=100,
                        sigma2=sigma2,
                        R=A,n_indi = 10,
                        verbose = F,
                        printevery = 1e5)

#simu = readRDS('output/simu_correlation_checkassumption_betahat.rds')
```

Coverage, adjusted for correlation
```{r}
simu$coverage_adj_hc0
mean(simu$coverage_adj_hc0,na.rm=TRUE)
```

Coverage, not adjusted for correlation
```{r}
simu$coverage_unadj_hc0
mean(simu$coverage_unadj_hc0)
```


```{r}
ploter.temp(simu)
```

Check normality 

```{r,fig.width=12,fig.height=12}
par(mfrow=c(4,4))
apply(simu$est_adj,2,function(z){qqnorm(z);qqline(z)})
```

Normal assumption is fine.


Check variance of $\hat p$, estimated proportion

```{r}
par(mfrow=c(1,1))
plot(apply(simu$est_adj,2,sd),type='b',ylim=c(0,0.06))
lines(apply(simu$se_adj_hc0,2,mean,na.rm=T),type='b',col=3)
legend('bottomright',c('true p_hat se','estimated'),lty=c(1,1),pch=c(1,1),col=c(1,3))
```

The variance seems underestimated.

Also check the variance of $\hat\beta$, estimated scaled proportion(before normalizing to sum to 1)

```{r}
par(mfrow=c(1,1))
plot(apply(simu$beta_hat,2,sd),type='b',ylim=c(10,30))
lines(apply(simu$beta_se_adj_hc0,2,mean,na.rm=TRUE),type='b',col=3)
legend('bottomright',c('true p_hat se','estimated'),lty=c(1,1),pch=c(1,1),col=c(1,3))
```

Look at the coverage of $\hat\beta$: also under-covered

```{r}
alpha = 0.05
ci_l = simu$beta_hat - qnorm(1-alpha/2)*simu$beta_se_adj_hc0
ci_r = simu$beta_hat + qnorm(1-alpha/2)*simu$beta_se_adj_hc0
coverage_adj_hc0 = (simu$true_betas>=ci_l) & (simu$true_betas<=ci_r)
coverage_adj_hc0=apply(coverage_adj_hc0,2,mean,na.rm=T)
coverage_adj_hc0
mean(coverage_adj_hc0,na.rm=T)
```

## Does this due to the plug-in beta hat?

Let's calculate the variance using true beta

```{r}
set.seed(12345)
simu_trueb = simu_corr_simple(ref,b.m,nreps=100,
                        sigma2=sigma2,
                        R=A,n_indi = 10,
                        verbose = F,
                        printevery = 1e5,
                        true.beta.for.Sigma = TRUE)


#simu_trueb = readRDS('output/simu_correlation_checkassumption_truebeta.rds')
```

Coverage, adjusted for correlation
```{r}
simu_trueb$coverage_adj_hc0
mean(simu_trueb$coverage_adj_hc0,na.rm=T)
```

Coverage, not adjusted for correlation
```{r}
simu_trueb$coverage_unadj_hc0
mean(simu_trueb$coverage_unadj_hc0)
```

Check variance of $\hat p$, estimated proportion

```{r}
par(mfrow=c(1,1))
plot(apply(simu_trueb$est_adj,2,sd),type='b',ylim=c(0,0.06))
lines(apply(simu_trueb$se_adj_hc0,2,mean,na.rm=T),type='b',col=3)
legend('bottomright',c('true p_hat se','estimated'),lty=c(1,1),pch=c(1,1),col=c(1,3))
```


Also check the variance of $\hat\beta$, estimated scaled proportion(before normalizing to sum to 1)

```{r}
par(mfrow=c(1,1))
plot(apply(simu_trueb$beta_hat,2,sd),type='b',ylim=c(10,30))
lines(apply(simu_trueb$beta_se_adj_hc0,2,mean,na.rm=T),type='b',col=3)
legend('bottomright',c('true p_hat se','estimated'),lty=c(1,1),pch=c(1,1),col=c(1,3))
```

Look at the coverage of $\hat\beta$: covered

```{r}
alpha = 0.05
ci_l = simu_trueb$beta_hat - qnorm(1-alpha/2)*simu_trueb$beta_se_adj_hc0
ci_r = simu_trueb$beta_hat + qnorm(1-alpha/2)*simu_trueb$beta_se_adj_hc0
coverage_adj_hc0 = (simu_trueb$true_betas>=ci_l) & (simu_trueb$true_betas<=ci_r)
coverage_adj_hc0=apply(coverage_adj_hc0,2,mean,na.rm=T)
coverage_adj_hc0
mean(coverage_adj_hc0)
```
